---
title: "pbp_data_creation"
output: html_document
date: "2024-12-04"
---

```{r}
# Load required libraries
library(dplyr)
library(nflverse)
library(tidyr)
library(purrr)
library(zoo)
library(car)
library(progress)

# Function to process a single year's data
process_year <- function(year) {
  message(paste("\nProcessing year:", year))
  
  message("Step 1/7: Loading data...")
# Load injuries for current and previous year
injuries_data <- load_injuries((year-2):year) %>%
  mutate(injury_date = as.Date(date_modified))

# Modified function to calculate days since last injury for a player
calc_days_since_injury <- function(player_id, current_date) {
  player_injuries <- injuries_data %>%
    filter(gsis_id == player_id,
           injury_date < current_date,
           !is.na(injury_date)) %>%
    arrange(desc(injury_date))
  
  if(nrow(player_injuries) > 0) {
    return(as.numeric(difftime(current_date, player_injuries$injury_date[1], units = "days")))
  }
  return(NA)
}
get_team_injury_counts <- function(team, game_date) {
  # Ensure date is in proper format
  game_date <- as.Date(game_date)
  
  # First filter by team, then do all the date filtering
  team_injuries <- injuries_data %>%
    filter(team == !!team) %>%
    mutate(injury_date = as.Date(date_modified))
  
  # Past year injuries - first group by injury type
  year_ago <- game_date - 365
  year_injuries <- team_injuries %>%
    filter(injury_date >= year_ago,
           injury_date < game_date) %>%
    count(practice_primary_injury) %>%
    pivot_wider(
      names_from = practice_primary_injury,
      names_prefix = "injury_count_year_",
      values_from = n,
      values_fill = 0
    )
  
  # Past week injuries
  week_ago <- game_date - 7
  week_injuries <- team_injuries %>%
    filter(injury_date >= week_ago,
           injury_date < game_date) %>%
    count(practice_primary_injury) %>%
    pivot_wider(
      names_from = practice_primary_injury,
      names_prefix = "injury_count_week_",
      values_from = n,
      values_fill = 0
    )
    
  # Past week status counts
  week_status <- team_injuries %>%
    filter(injury_date >= week_ago,
           injury_date < game_date) %>%
    summarise(
      week_questionable = sum(report_status == "Questionable", na.rm = TRUE),
      week_doubtful = sum(report_status == "Doubtful", na.rm = TRUE),
      week_out = sum(report_status == "Out", na.rm = TRUE),
      week_limited_practice = sum(practice_status == "Limited Participation in Practice", na.rm = TRUE),
      week_dnp = sum(practice_status == "Did Not Participate In Practice", na.rm = TRUE)
    )
  
  # Combine all counts
  bind_cols(year_injuries, week_injuries, week_status)
}

  
  # Load play by play data and participation data for the year
  play_year <- load_pbp(year)
  participation_data <- load_participation(year) %>%
    select(
      nflverse_game_id,
      play_id,
      offense_formation,
      offense_personnel,
      defense_personnel,
      defenders_in_box,
      number_of_pass_rushers,
      offense_players,
      defense_players,
      defense_man_zone_type,
      defense_coverage_type
    )
  
  # Load roster data for player information
  roster_data <- load_rosters(year)
  
    message("Step 2/7: Processing weekly stats...")
  # Calculate weekly run stats
  weekly_run_stats <- play_year %>%
    filter(play_type %in% c("pass", "run"), down %in% 1:3) %>%
    group_by(posteam, week, down) %>%
    summarize(
      weekly_run_plays = sum(play_type == "run", na.rm = TRUE),
      weekly_total_plays = n(),
      .groups = 'drop'
    ) %>%
    arrange(posteam, week, down) %>%
    group_by(posteam, down) %>%
    mutate(
      cum_run_plays = cumsum(weekly_run_plays),
      cum_total_plays = cumsum(weekly_total_plays)
    ) %>%
    ungroup() %>%
    pivot_wider(
      id_cols = c(posteam, week),
      names_from = down,
      names_glue = "down{down}_{.value}",
      values_from = c(cum_run_plays, cum_total_plays)
    ) %>%
    arrange(posteam, week)

  # Calculate run percentages
  run_pcts <- weekly_run_stats %>%
    mutate(
      down1_pct = round(100 * down1_cum_run_plays / down1_cum_total_plays, 1),
      down2_pct = round(100 * down2_cum_run_plays / down2_cum_total_plays, 1),
      down3_pct = round(100 * down3_cum_run_plays / down3_cum_total_plays, 1)
    ) %>%
    select(posteam, week, down1_pct, down2_pct, down3_pct)

  # Calculate prep time
  prep_time_df <- play_year %>%
    select(game_id, game_date, posteam = home_team) %>%
    distinct() %>%
    bind_rows(
      play_year %>%
        select(game_id, game_date, posteam = away_team) %>%
        distinct()
    ) %>%
    arrange(posteam, game_date) %>%
    group_by(posteam) %>%
    mutate(
      last_game_date = lag(game_date),
      prep_days = as.numeric(difftime(game_date, last_game_date, units = "days"))
    ) %>%
    ungroup()
  
add_player_info <- function(df, players_data, player_col, prefix) {
    if (!player_col %in% names(df)) {
        return(df)
    }
    
    new_cols <- c("full_name", "birth_date", "rookie_year", "draft_number", "jersey_number", "pff_id", 
                "weight", "height", "position", "depth_chart_position")
    
    # Create NA columns
    for(col in new_cols) {
        new_col_name <- paste0(prefix, "_", col)
        df[[new_col_name]] <- NA
    }
    
    # Add injury history columns
    df[[paste0(prefix, "_days_since_injury")]] <- NA
    df[[paste0(prefix, "_last_injury_type")]] <- NA
    df[[paste0(prefix, "_injury_count_1yr")]] <- NA
    
    # Add practice/report status columns for past week
    df[[paste0(prefix, "_week_questionable")]] <- 0
    df[[paste0(prefix, "_week_doubtful")]] <- 0
    df[[paste0(prefix, "_week_out")]] <- 0
    df[[paste0(prefix, "_week_limited_practice")]] <- 0
    df[[paste0(prefix, "_week_dnp")]] <- 0
    df[[paste0(prefix, "_current_status")]] <- NA  # Most recent status
    df[[paste0(prefix, "_current_practice")]] <- NA  # Most recent practice status
    
    # Get valid players
    valid_players <- unique(df[[player_col]][!is.na(df[[player_col]])])
    
    if(length(valid_players) > 0) {
        # Get player info and handle duplicates
        players_info <- players_data %>%
            filter(gsis_id %in% valid_players) %>%
            group_by(gsis_id) %>%
            slice_tail(n = 1) %>%
            ungroup() %>%
            select(gsis_id, all_of(new_cols))
        
        # Update player information
        for(i in 1:nrow(df)) {
            if(!is.na(df[[player_col]][i])) {
                player_id <- df[[player_col]][i]
                current_date <- as.Date(df$game_date[i])
                player_data <- players_info %>% filter(gsis_id == player_id)
                
                # Update basic player info
                if(nrow(player_data) == 1) {
                    for(col in new_cols) {
                        if(!is.null(player_data[[col]])) {
                            df[i, paste0(prefix, "_", col)] <- player_data[[col]][1]
                        }
                    }
                }
                
                # Get player's injury history
                player_injuries <- injuries_data %>%
                    filter(gsis_id == player_id,
                           injury_date < current_date,
                           !is.na(injury_date)) %>%
                    arrange(desc(injury_date))
                
                # Update injury info
                if(nrow(player_injuries) > 0) {
                    # Days since last injury
                    df[i, paste0(prefix, "_days_since_injury")] <- 
                        as.numeric(difftime(current_date, player_injuries$injury_date[1], units = "days"))
                    
                    # Last injury type
                    df[i, paste0(prefix, "_last_injury_type")] <- 
                        player_injuries$practice_primary_injury[1]
                    
                    # Count of injuries in past year
                    one_year_ago <- current_date - 365
                    df[i, paste0(prefix, "_injury_count_1yr")] <- 
                        sum(player_injuries$injury_date >= one_year_ago)
                    
                    # Get past week's statuses
                    week_ago <- current_date - 7
                    recent_injuries <- injuries_data %>%
                        filter(gsis_id == player_id,
                               injury_date >= week_ago,
                               injury_date < current_date)
                    
                    if(nrow(recent_injuries) > 0) {
                        # Count statuses in past week
                        df[i, paste0(prefix, "_week_questionable")] <- 
                            sum(recent_injuries$report_status == "Questionable", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_doubtful")] <- 
                            sum(recent_injuries$report_status == "Doubtful", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_out")] <- 
                            sum(recent_injuries$report_status == "Out", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_limited_practice")] <- 
                            sum(recent_injuries$practice_status == "Limited Participation in Practice", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_dnp")] <- 
                            sum(recent_injuries$practice_status == "Did Not Participate In Practice", na.rm = TRUE)
                        
                        # Most recent status
                        df[i, paste0(prefix, "_current_status")] <- recent_injuries$report_status[1]
                        df[i, paste0(prefix, "_current_practice")] <- recent_injuries$practice_status[1]
                    }
                }
            }
        }
    }
    return(df)
}
  message("Step 3/7: Processing third and fourth downs...")

  # Process third downs
  third_downs <- play_year %>%
    filter(down == 3) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      season = year
    ) %>%
    filter(!field_goal_attempt, !punt_attempt, play_type %in% c("pass", "run")) %>%
    mutate(converted = first_down) %>%
    select(-first_down) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id")) %>%
    separate(offense_players, 
             into = paste0("offense_player_", 1:11), 
             sep = ";", 
             fill = "right") %>%
    separate(defense_players, 
             into = paste0("defense_player_", 1:11), 
             sep = ";", 
             fill = "right")

  # Process fourth downs
  fourth_downs <- play_year %>%
    filter(down == 4) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      attempt = if_else(play_type %in% c("punt", "field_goal"), 0, 1),
      minutes_remaining = game_seconds_remaining / 60,
      qtr = ceiling(minutes_remaining / 15),
      season = year
    ) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id")) %>%
    separate(offense_players, 
             into = paste0("offense_player_", 1:11), 
             sep = ";", 
             fill = "right") %>%
    separate(defense_players, 
             into = paste0("defense_player_", 1:11), 
             sep = ";", 
             fill = "right")
  
  message("Step 4/7: Adding player information...")
  # Add player information for both offensive and defensive players
  for(i in 1:11) {
    # Process offensive players
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("offense_player_", i), 
                                  paste0("offense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("offense_player_", i), 
                                   paste0("offense_player_", i))
    
    # Process defensive players
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("defense_player_", i), 
                                  paste0("defense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("defense_player_", i), 
                                   paste0("defense_player_", i))
  }
    message("Step 5/7: Processing team injury data...")
    
    # Add defteam and join with run percentages and prep time
  third_downs <- third_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  fourth_downs <- fourth_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  # Inside process_year function, after adding individual player injuries
third_downs <- third_downs %>%
  rowwise() %>%
  mutate(
    offense_injuries = list(get_team_injury_counts(posteam, game_date)),
    defense_injuries = list(get_team_injury_counts(defteam, game_date))
  ) %>%
  ungroup() %>%
  unnest_wider(c(offense_injuries, defense_injuries), 
               names_sep = "_")

fourth_downs <- fourth_downs %>%
  rowwise() %>%
  mutate(
    offense_injuries = list(get_team_injury_counts(posteam, game_date)),
    defense_injuries = list(get_team_injury_counts(defteam, game_date))
  ) %>%
  ungroup() %>%
  unnest_wider(c(offense_injuries, defense_injuries), 
               names_sep = "_")

# Add team-level injury aggregates
third_downs <- third_downs %>%
  group_by(posteam, game_date) %>%
  mutate(
    offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
    offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
  ) %>%
  group_by(defteam, game_date) %>%
  mutate(
    defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
    defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
  ) %>%
  ungroup()

fourth_downs <- fourth_downs %>%
  group_by(posteam, game_date) %>%
  mutate(
    offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
    offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
  ) %>%
  group_by(defteam, game_date) %>%
  mutate(
    defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
    defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
  ) %>%
  ungroup()

  

  # Calculate previous game stats
  all_teams_weeks <- bind_rows(
    select(third_downs, defteam, week),
    select(fourth_downs, defteam, week)
  ) %>% distinct()

  prev_stats <- map2_dfr(
    all_teams_weeks$defteam,
    all_teams_weeks$week,
    ~{
      prev_game <- play_year %>%
        filter((home_team == .x | away_team == .x) & week < .y) %>%
        arrange(desc(week)) %>%
        slice(1) %>%
        pull(week)
      
      if(length(prev_game) == 0) return(tibble(
        defteam = .x,
        week = .y,
        attempts = 0,
        successes = 0,
        failures = 0
      ))
      
      play_year %>%
        filter(week == prev_game,
               (home_team == .x | away_team == .x),
               down == 4,
               !play_type %in% c("punt", "field_goal")) %>%
        summarize(
          defteam = .x,
          week = .y,
          attempts = n(),
          successes = sum(first_down == 1, na.rm = TRUE),
          failures = sum(first_down == 0, na.rm = TRUE)
        )
    })
  message("Step 6/7: Processing model data...")

  # Process final model data
  model_3rd <- third_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
    left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  model_4th <- fourth_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
    left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()
  
  message("Step 7/7: Adding specialists...")

  # Add specialists
  rd2 <- load_rosters_weekly(year)
  specialists <- rd2 %>%
    filter(position %in% c("K", "P")) %>%
    select(week, team, full_name, position, gsis_id)

  model_4th <- model_4th %>%
    left_join(specialists %>% 
                filter(position == "K") %>% 
                select(week, team, k_name=full_name, k_gsis_id=gsis_id),
              by=c("week", "posteam"="team")) %>%
    left_join(specialists %>% 
                filter(position == "P") %>% 
                select(week, team, p_name=full_name, p_gsis_id=gsis_id),
              by=c("week", "posteam"="team"))

  model_3rd <- model_3rd %>%
    left_join(specialists %>% 
                filter(position == "K") %>% 
                select(week, team, k_name=full_name, k_gsis_id=gsis_id),
              by=c("week", "posteam"="team")) %>%
    left_join(specialists %>% 
                filter(position == "P") %>% 
                select(week, team, p_name=full_name, p_gsis_id=gsis_id),
              by=c("week", "posteam"="team"))

  return(list(
    third_downs = model_3rd,
    fourth_downs = model_4th
  ))
}

# Process all years
```

```{r}
# Process all years
years <- 2016:2023
all_data <- list()

# Process each year
for(year in years) {
  all_data[[as.character(year)]] <- process_year(year)
}

# Combine all years
combined_third_downs <- bind_rows(
  lapply(all_data, function(x) x$third_downs)
)

combined_fourth_downs <- bind_rows(
  lapply(all_data, function(x) x$fourth_downs)
)
```

kill minutes_remaining and qtr

```{r}
combined_fourth_downs <- combined_fourth_downs %>%
  select(-minutes_remaining, -qtr)
```






```{r}
write.csv(combined_third_downs, "ctd.csv", row.names = FALSE)
write.csv(combined_fourth_downs, "cfd.csv", row.names = FALSE)
```


# Explinations and proofs

offense_player_1_days_since_injury       # Days since last injury for player 

offense_player_1_current_status         # Most recent game status (Questionable/Doubtful/Out/NA)
offense_player_1_current_practice       # Most recent practice status (Limited/DNP/Full/NA)

these look 1 week back

offense_player_1_week_questionable      # Number of times listed as "Questionable" on injury report
offense_player_1_week_doubtful          # Number of times listed as "Doubtful" on injury report  
offense_player_1_week_out               # Number of times listed as "Out" on injury report
offense_player_1_week_limited_practice  # Number of times had "Limited Participation in Practice" 
offense_player_1_week_dnp               # Number of times "Did Not Participate in Practice"


```{r}
library(dplyr)
library(nflverse)

# Load multiple years of injury data
injury_data <- load_injuries(2016:2023)

# Get unique injury types and sort them
injury_types <- injury_data %>%
  select(practice_primary_injury) %>%
  distinct() %>%
  filter(!is.na(practice_primary_injury)) %>%
  arrange(practice_primary_injury)

print("All unique injury types in data:")
print(injury_types)

# Get frequency of each injury type
injury_frequencies <- injury_data %>%
  group_by(practice_primary_injury) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  filter(!is.na(practice_primary_injury))

print("\nInjury types by frequency:")
print(injury_frequencies)
```


offense_injuries_injury_count_year   # Count of knee injuries for offense in past 365 days
offense_injuries_injury_count_week   # Count of knee injuries for offense in past 7 days
defense_injuries_injury_count_year   # Count of knee injuries for defense in past 365 days
defense_injuries_injury_count_week   # Count of knee injuries for defense in past 7 days


Proof that the injuries are proper
```{r}
t1 <- combined_fourth_downs[5000, ]
```


```{r}
t1$game_id
# date of game
t1$game_date
```

game date
```{r}
t1$game_date
```

full names of all players on the feild
```{r}
t1$offense_player_1_full_name
t1$defense_player_1_full_name
```

```{r}
t1$offense_player_1_days_since_injury
t1$defense_player_1_days_since_injury
```
load injury data

```{r}
inj1 <- load_injuries(2017)
```


get the injuries of players with the same names as either offensive of defensive players

```{r}
filt_inj <- inj1 %>% filter(full_name %in% c(t1$offense_player_1_full_name, t1$defense_player_1_full_name))
```

select the dates of the injuries for each player 
```{r}
filt_inj %>% select(full_name, date_modified)
```
the game was on the 10th of september 2017. we can see a injury on the 6th. Our 4 days since injury is accurate.


injury count for the year
```{r}
inj2 <- load_injuries(2016:2017)
```

```{r}

# get Deandre hopkins injury data from 2017 and 2016 
inj2 <- inj2 %>% filter(full_name ==  t1$offense_player_1_full_name)
```


get the injuries that happened before the date of the game
```{r}
inj2 <- inj2 %>% filter(date_modified < t1$game_date)
```

```{r}
nrow(inj2)
t1$offense_player_1_injury_count_1yr
```

offense_injury_count_year_Ankle # Count of ankle injuries for offense in past 365 days
offense_injury_count_week_Ankle # Count of ankle injuries for offense in past 7 days



```{r}
# Convert game_date to Date object
game_date <- as.Date(t1$game_date)

# Test for week count
inj3_week <- inj3 %>% 
  mutate(date_modified = as.Date(date_modified)) %>%
  filter(date_modified >= game_date - 7,
         date_modified < game_date)

# Test for year count
inj3_year <- inj3 %>%
  mutate(date_modified = as.Date(date_modified)) %>%
  filter(date_modified >= game_date - 365,
         date_modified < game_date)

print("Week count:")
print(nrow(inj3_week))
print("Year count:")
print(nrow(inj3_year))

# Also print actual values we're comparing against from t1
print("Values from dataset:")
print(paste("Week count in data:", t1$offense_injuries_injury_count_week_Ankle))
print(paste("Year count in data:", t1$offense_injuries_injury_count_year_Ankle))
```
