---
title: "pbp_data_creation"
output: html_document
date: "2024-12-04"
---



```{r}
# Load required libraries
library(dplyr)
library(nflverse)
library(tidyr)
library(purrr)
library(zoo)
library(car)
library(progress)

# Validation function
validate_data <- function(df) {
  dupes <- duplicated(names(df))
  if(any(dupes)) {
    warning("Duplicate columns found: ", paste(names(df)[dupes], collapse=", "))
  }
  
  req_cols <- c("play_id", "game_id", "posteam")
  missing <- req_cols[!req_cols %in% names(df)]
  if(length(missing) > 0) {
    stop("Missing required columns: ", paste(missing, collapse=", "))
  }
  return(df)
}


# Get specialists function
get_specialists <- function(year) {
  pbp <- bind_rows(
    load_pbp(year),
    load_pbp(year - 1)
  ) %>%
  arrange(game_date, desc(play_id))
  
  kickers <- pbp %>%
    filter(field_goal_attempt == 1 | extra_point_attempt == 1) %>%
    select(team = posteam, week, season, kicker_player_id, kicker_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  punters <- pbp %>%
    filter(punt_attempt == 1) %>%
    select(team = posteam, week, season, punter_player_id, punter_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  list(kickers = kickers, punters = punters)
}
add_player_info <- function(df, players_data, player_col, prefix) {
    if (!player_col %in% names(df)) {
        return(df)
    }
    
    new_cols <- c("full_name", "birth_date", "rookie_year", "draft_number", "jersey_number", "pff_id", 
                "weight", "height", "position", "depth_chart_position")
    
    for(col in new_cols) {
        new_col_name <- paste0(prefix, "_", col)
        df[[new_col_name]] <- NA
    }
    
    df[[paste0(prefix, "_days_since_injury")]] <- NA
    df[[paste0(prefix, "_last_injury_type")]] <- NA
    df[[paste0(prefix, "_injury_count_1yr")]] <- NA
    
    df[[paste0(prefix, "_week_questionable")]] <- 0
    df[[paste0(prefix, "_week_doubtful")]] <- 0
    df[[paste0(prefix, "_week_out")]] <- 0
    df[[paste0(prefix, "_week_limited_practice")]] <- 0
    df[[paste0(prefix, "_week_dnp")]] <- 0
    df[[paste0(prefix, "_current_status")]] <- NA
    df[[paste0(prefix, "_current_practice")]] <- NA
    
    valid_players <- unique(df[[player_col]][!is.na(df[[player_col]])])
    
    if(length(valid_players) > 0) {
        players_info <- players_data %>%
            filter(gsis_id %in% valid_players) %>%
            group_by(gsis_id) %>%
            slice_tail(n = 1) %>%
            ungroup() %>%
            select(gsis_id, all_of(new_cols))
        
        for(i in 1:nrow(df)) {
            if(!is.na(df[[player_col]][i])) {
                player_id <- df[[player_col]][i]
                current_date <- as.Date(df$game_date[i])
                player_data <- players_info %>% filter(gsis_id == player_id)
                
                if(nrow(player_data) == 1) {
                    for(col in new_cols) {
                        if(!is.null(player_data[[col]])) {
                            df[i, paste0(prefix, "_", col)] <- player_data[[col]][1]
                        }
                    }
                }
                
                player_injuries <- injuries_data %>%
                    filter(gsis_id == player_id,
                           injury_date < current_date,
                           !is.na(injury_date)) %>%
                    arrange(desc(injury_date))
                
                if(nrow(player_injuries) > 0) {
                    df[i, paste0(prefix, "_days_since_injury")] <- 
                        as.numeric(difftime(current_date, player_injuries$injury_date[1], units = "days"))
                    
                    df[i, paste0(prefix, "_last_injury_type")] <- 
                        player_injuries$practice_primary_injury[1]
                    
                    one_year_ago <- current_date - 365
                    df[i, paste0(prefix, "_injury_count_1yr")] <- 
                        sum(player_injuries$injury_date >= one_year_ago)
                    
                    week_ago <- current_date - 7
                    recent_injuries <- injuries_data %>%
                        filter(gsis_id == player_id,
                               injury_date >= week_ago,
                               injury_date < current_date)
                    
                    if(nrow(recent_injuries) > 0) {
                        df[i, paste0(prefix, "_week_questionable")] <- 
                            sum(recent_injuries$report_status == "Questionable", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_doubtful")] <- 
                            sum(recent_injuries$report_status == "Doubtful", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_out")] <- 
                            sum(recent_injuries$report_status == "Out", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_limited_practice")] <- 
                            sum(recent_injuries$practice_status == "Limited Participation in Practice", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_dnp")] <- 
                            sum(recent_injuries$practice_status == "Did Not Participate In Practice", na.rm = TRUE)
                        
                        df[i, paste0(prefix, "_current_status")] <- recent_injuries$report_status[1]
                        df[i, paste0(prefix, "_current_practice")] <- recent_injuries$practice_status[1]
                    }
                }
            }
        }
    }
    return(df)
}
```


```{r}
process_year <- function(year) {
 message(paste("\nProcessing year:", year))
  
  message("Step 1/7: Loading data...")
  # Add global assignment
  injuries_data <<- load_injuries((year-2):year) %>%
    mutate(injury_date = as.Date(date_modified))

  top_injury_types <- injuries_data %>%
    filter(!is.na(practice_primary_injury)) %>%
    count(practice_primary_injury, sort = TRUE) %>%
    slice_head(n = 9) %>%
    pull(practice_primary_injury)

  get_team_injury_counts <- function(team, game_date) {
    game_date <- as.Date(game_date)
    
    team_injuries <- injuries_data %>%
      filter(team == !!team) %>%
      mutate(
        injury_date = as.Date(date_modified),
        standard_injury = case_when(
          practice_primary_injury %in% top_injury_types ~ practice_primary_injury,
          TRUE ~ "Other"
        )
      )
    
    year_ago <- game_date - 365
    year_injuries <- team_injuries %>%
      filter(injury_date >= year_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_year_",
        values_from = n,
        values_fill = 0
      )
    
    week_ago <- game_date - 7
    week_injuries <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_week_",
        values_from = n,
        values_fill = 0
      )
      
    week_status <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      summarise(
        week_questionable = sum(report_status == "Questionable", na.rm = TRUE),
        week_doubtful = sum(report_status == "Doubtful", na.rm = TRUE),
        week_out = sum(report_status == "Out", na.rm = TRUE),
        week_limited_practice = sum(practice_status == "Limited Participation in Practice", na.rm = TRUE),
        week_dnp = sum(practice_status == "Did Not Participate In Practice", na.rm = TRUE)
      )
    
    bind_cols(year_injuries, week_injuries, week_status)
  }

  play_year <- load_pbp(year)
  participation_data <- load_participation(year) %>%
    select(
      nflverse_game_id,
      play_id,
      offense_formation,
      offense_personnel,
      defense_personnel,
      defenders_in_box,
      number_of_pass_rushers,
      offense_players,
      defense_players,
      defense_man_zone_type,
      defense_coverage_type
    )
    
  roster_data <- load_rosters(year)
    
  message("Step 2/7: Processing weekly stats...")
  weekly_run_stats <- play_year %>%
    filter(play_type %in% c("pass", "run"), down %in% 1:3) %>%
    group_by(posteam, week, down) %>%
    summarize(
      weekly_run_plays = sum(play_type == "run", na.rm = TRUE),
      weekly_total_plays = n(),
      .groups = 'drop'
    ) %>%
    arrange(posteam, week, down) %>%
    group_by(posteam, down) %>%
    mutate(
      cum_run_plays = cumsum(weekly_run_plays),
      cum_total_plays = cumsum(weekly_total_plays)
    ) %>%
    ungroup() %>%
    pivot_wider(
      id_cols = c(posteam, week),
      names_from = down,
      names_glue = "down{down}_{.value}",
      values_from = c(cum_run_plays, cum_total_plays)
    ) %>%
    arrange(posteam, week)

  run_pcts <- weekly_run_stats %>%
    mutate(
      down1_pct = round(100 * down1_cum_run_plays / down1_cum_total_plays, 1),
      down2_pct = round(100 * down2_cum_run_plays / down2_cum_total_plays, 1),
      down3_pct = round(100 * down3_cum_run_plays / down3_cum_total_plays, 1)
    ) %>%
    select(posteam, week, down1_pct, down2_pct, down3_pct)

  prep_time_df <- play_year %>%
    select(game_id, game_date, posteam = home_team) %>%
    distinct() %>%
    bind_rows(
      play_year %>%
        select(game_id, game_date, posteam = away_team) %>%
        distinct()
    ) %>%
    arrange(posteam, game_date) %>%
    group_by(posteam) %>%
    mutate(
      last_game_date = lag(game_date),
      prep_days = as.numeric(difftime(game_date, last_game_date, units = "days"))
    ) %>%
    ungroup()

  # Add after run_pcts but before prep_time_df in Step 2/7:
  message("Step 2.5/7: Processing coach history...")
  
  # Get coach records across all games
  coach_records <- play_year %>%
    select(game_id, game_date, home_team, away_team, home_coach, away_coach,
           home_score, away_score) %>%
    distinct() %>%
    # Get home coach records
    mutate(
      home_win = home_score > away_score,
      home_loss = home_score < away_score
    ) %>%
    group_by(home_coach) %>%
    summarize(
      overall_wins = sum(home_win, na.rm = TRUE),
      overall_losses = sum(home_loss, na.rm = TRUE),
      overall_win_pct = round(sum(home_win, na.rm = TRUE) / n(), 3)
    ) %>%
    # Combine with away coach records
    bind_rows(
      play_year %>%
      select(game_id, away_coach, home_score, away_score) %>%
      distinct() %>%
      group_by(away_coach) %>%
      summarize(
        overall_wins = sum(away_score > home_score, na.rm = TRUE),
        overall_losses = sum(away_score < home_score, na.rm = TRUE),
        overall_win_pct = round(sum(away_score > home_score, na.rm = TRUE) / n(), 3)
      )
    ) %>%
    group_by(home_coach) %>%
    summarize(
      overall_wins = sum(overall_wins),
      overall_losses = sum(overall_losses),
      overall_win_pct = round(mean(overall_win_pct), 3)
    ) %>%
    rename(coach = home_coach)

  # Get coach records with current team only
  current_team_records <- play_year %>%
    select(game_id, game_date, posteam, home_team, away_team, home_coach, away_coach,
           posteam_score, defteam_score) %>%
    distinct() %>%
    mutate(
      coach = ifelse(posteam == home_team, home_coach, away_coach)
    ) %>%
    group_by(coach, posteam) %>%
    summarize(
      team_wins = sum(posteam_score > defteam_score, na.rm = TRUE),
      team_losses = sum(posteam_score < defteam_score, na.rm = TRUE),
      team_win_pct = round(mean(posteam_score > defteam_score, na.rm = TRUE), 3),
      .groups = 'drop'
    )
  
  # Add after coach records calculation in Step 2.5:
message("Diagnosing coach records duplicates...")
coach_dupes <- coach_records %>%
  group_by(coach) %>%
  filter(n() > 1)

print("Number of duplicate coach record rows:")
print(nrow(coach_dupes))
if(nrow(coach_dupes) > 0) {
  print("Sample of duplicates:")
  print(head(coach_dupes))
}
  
 message("Step 2.6/7: Processing coach tenure...")
# Calculate coach tenure - modify this section
coach_tenure <- play_year %>%
    select(game_date, home_team, away_team, home_coach, away_coach) %>%
    # Get unique games first
    distinct() %>%
    pivot_longer(
      cols = c(home_coach, away_coach),
      names_to = "role",
      values_to = "coach"
    ) %>%
    mutate(team = if_else(role == "home_coach", home_team, away_team)) %>%
    # Get first appearance for each coach-team combo
    group_by(coach, team) %>%
    summarize(
      first_appearance = min(game_date),
      .groups = 'drop'
    )

# Calculate tenure for each game - modify this section
game_tenure <- play_year %>%
    select(game_id, game_date, posteam, home_team, away_team, home_coach, away_coach) %>%
    # Get unique game-team combinations
    distinct() %>%
    mutate(
      coach = ifelse(posteam == home_team, home_coach, away_coach)
    ) %>%
    left_join(coach_tenure, by = c("coach", "posteam" = "team")) %>%
    mutate(
      tenure_days = as.numeric(as.Date(game_date) - as.Date(first_appearance))
    )
  
  # Add after game tenure calculation in Step 2.6:
message("Diagnosing game tenure duplicates...")
tenure_dupes <- game_tenure %>%
  group_by(game_id, posteam, coach) %>%
  filter(n() > 1)

print("Number of duplicate tenure rows:")
print(nrow(tenure_dupes))
if(nrow(tenure_dupes) > 0) {
  print("Sample of duplicates:")
  print(head(tenure_dupes))
}

# Add after the game tenure diagnostic check:
if(nrow(tenure_dupes) > 0) {
    print("Detailed look at tenure duplicates:")
    print(
        tenure_dupes %>%
        arrange(game_id, posteam, coach) %>%
        select(game_id, game_date, posteam, coach, first_appearance, tenure_days) %>%
        head(10)
    )
    
    # Count frequency of duplications
    print("Number of duplicates per game/team/coach combination:")
    print(
        tenure_dupes %>%
        group_by(game_id, posteam, coach) %>%
        summarise(n = n(), .groups = 'drop') %>%
        count(n) %>%
        arrange(desc(n))
    )
}


  
  message("Step 2.7/7: Processing formation and defensive tendencies...")
  # Formation tendencies by down
  formation_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(posteam, down) %>%
    summarize(
      total_formation_plays = sum(!is.na(offense_formation)),
      
      # Formation usage rates
      shotgun_rate = ifelse(total_formation_plays > 0,
                           sum(offense_formation == "SHOTGUN", na.rm = TRUE) / total_formation_plays, 0),
      singleback_rate = ifelse(total_formation_plays > 0,
                              sum(offense_formation == "SINGLEBACK", na.rm = TRUE) / total_formation_plays, 0),
      empty_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "EMPTY", na.rm = TRUE) / total_formation_plays, 0),
      iform_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "I_FORM", na.rm = TRUE) / total_formation_plays, 0),
      pistol_rate = ifelse(total_formation_plays > 0,
                          sum(offense_formation == "PISTOL", na.rm = TRUE) / total_formation_plays, 0),
      wildcat_rate = ifelse(total_formation_plays > 0,
                           sum(offense_formation == "WILDCAT", na.rm = TRUE) / total_formation_plays, 0),
      jumbo_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "JUMBO", na.rm = TRUE) / total_formation_plays, 0),
      
      # Calculate success rates for each formation
      shotgun_success = mean(first_down[offense_formation == "SHOTGUN"], na.rm = TRUE),
      singleback_success = mean(first_down[offense_formation == "SINGLEBACK"], na.rm = TRUE),
      empty_success = mean(first_down[offense_formation == "EMPTY"], na.rm = TRUE),
      iform_success = mean(first_down[offense_formation == "I_FORM"], na.rm = TRUE),
      pistol_success = mean(first_down[offense_formation == "PISTOL"], na.rm = TRUE),
      wildcat_success = mean(first_down[offense_formation == "WILDCAT"], na.rm = TRUE),
      jumbo_success = mean(first_down[offense_formation == "JUMBO"], na.rm = TRUE),
      .groups = 'drop'
    )

  # Defense tendencies by down
  defense_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(defteam, down) %>%
    summarize(
      avg_pass_rushers = mean(number_of_pass_rushers, na.rm = TRUE),
      avg_box_defenders = mean(defenders_in_box, na.rm = TRUE),
      
      # Coverage type rates - using all known types
      total_coverage_plays = sum(!is.na(defense_coverage_type)),
      cover0_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_0", na.rm = TRUE) / total_coverage_plays, 0),
      cover1_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_1", na.rm = TRUE) / total_coverage_plays, 0),
      cover2_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_2", na.rm = TRUE) / total_coverage_plays, 0),
      cover3_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_3", na.rm = TRUE) / total_coverage_plays, 0),
      cover4_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_4", na.rm = TRUE) / total_coverage_plays, 0),
      cover6_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_6", na.rm = TRUE) / total_coverage_plays, 0),
      two_man_rate = ifelse(total_coverage_plays > 0, 
                           sum(defense_coverage_type == "2_MAN", na.rm = TRUE) / total_coverage_plays, 0),
      prevent_rate = ifelse(total_coverage_plays > 0, 
                           sum(defense_coverage_type == "PREVENT", na.rm = TRUE) / total_coverage_plays, 0),
      .groups = 'drop'
    )

  # Coverage tendencies by down
  coverage_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(defteam, down, defense_coverage_type) %>%
    summarize(
      coverage_plays = n(),
      coverage_success_rate = mean(first_down == 0, na.rm = TRUE),  # defense success = stop
      .groups = 'drop'
    ) %>%
    group_by(defteam, down) %>%
    mutate(
      coverage_usage_rate = coverage_plays / sum(coverage_plays)
    ) %>%
    ungroup()
  
  
message("Step 3/7: Processing third and fourth downs...")
  third_downs <- play_year %>%
    filter(down == 3) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      season = year
    ) %>%
    filter(!field_goal_attempt, !punt_attempt, play_type %in% c("pass", "run")) %>%
    mutate(converted = first_down) %>%
    select(-first_down) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id"))

  fourth_downs <- play_year %>%
    filter(down == 4) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      attempt = if_else(play_type %in% c("punt", "field_goal"), 0, 1),
      season = year
    ) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id"))

  message("Step 3.5/7: Assigning players to positions...")
  
  # Create empty player columns for third downs
  message("  Creating player columns for third downs...")
  for(i in 1:11) {
    third_downs[[paste0("offense_player_", i)]] <- NA_character_
    third_downs[[paste0("defense_player_", i)]] <- NA_character_
  }
  
  # Create empty player columns for fourth downs
  message("  Creating player columns for fourth downs...")
  for(i in 1:11) {
    fourth_downs[[paste0("offense_player_", i)]] <- NA_character_
    fourth_downs[[paste0("defense_player_", i)]] <- NA_character_
  }
  
  # Function to assign players to positions
  assign_players_to_positions <- function(players_data, roster_data, is_offense = TRUE) {
    # Initialize empty slots
    player_slots <- rep(NA_character_, 11)
    
    # Skip if empty player data
    if(is.null(players_data) || players_data == "") {
      return(player_slots)
    }
    
    # Get player ids from semicolon-separated string
    player_ids <- unlist(strsplit(players_data, ";"))
    if(length(player_ids) == 0) {
      return(player_slots)
    }
    
    # Get player positions from roster data
    players_info <- roster_data %>%
      filter(gsis_id %in% player_ids) %>%
      select(gsis_id, depth_chart_position)
    
    if(nrow(players_info) == 0) {
      return(player_slots)
    }
    
    # First check for special teams
    st_players <- players_info %>%
      filter(depth_chart_position %in% c("P", "K", "LS"))
    if(nrow(st_players) > 0) {
      player_slots[1:min(11, nrow(st_players))] <- st_players$gsis_id[1:min(11, nrow(st_players))]
      return(player_slots)
    }
    
    # If offense
    if(is_offense) {
      # QB first
      qbs <- players_info %>% filter(depth_chart_position == "QB")
      if(nrow(qbs) > 0) player_slots[1] <- qbs$gsis_id[1]
      
      # RB/FB next
      backs <- players_info %>% filter(depth_chart_position %in% c("RB", "FB"))
      if(nrow(backs) > 0) {
        for(i in 1:min(2, nrow(backs))) {
          player_slots[1 + i] <- backs$gsis_id[i]
        }
      }
      
      # WR next
      wrs <- players_info %>% filter(depth_chart_position == "WR")
      wr_start <- if(nrow(backs) >= 2) 4 else 3
      if(nrow(wrs) > 0 && wr_start <= 11) {
        for(i in 1:min(11 - wr_start + 1, nrow(wrs))) {
          player_slots[wr_start + i - 1] <- wrs$gsis_id[i]
        }
      }
      
      # Find next empty slot
      next_empty <- which(is.na(player_slots))
      if(length(next_empty) > 0) {
        next_slot <- min(next_empty)
        
        # TE next
        tes <- players_info %>% filter(depth_chart_position == "TE")
        if(nrow(tes) > 0 && next_slot <= 11) {
          player_slots[next_slot] <- tes$gsis_id[1]
          next_slot <- next_slot + 1
        }
        
        # O-line last
        oline <- players_info %>% 
          filter(depth_chart_position %in% c("T", "G", "C", "OT", "OG"))
        if(nrow(oline) > 0 && next_slot <= 11) {
          for(i in 1:min(11 - next_slot + 1, nrow(oline))) {
            player_slots[next_slot + i - 1] <- oline$gsis_id[i]
          }
        }
      }
    } else {
      # Defense
      # D-line first
      dline <- players_info %>% 
        filter(depth_chart_position %in% c("DE", "DT", "NT", "DL"))
      if(nrow(dline) > 0) {
        for(i in 1:min(3, nrow(dline))) {
          player_slots[i] <- dline$gsis_id[i]
        }
      }
      
      # Find next empty slot
      next_empty <- which(is.na(player_slots))
      if(length(next_empty) > 0) {
        next_slot <- min(next_empty)
        
        # Linebackers next
        lbs <- players_info %>% 
          filter(depth_chart_position %in% c("MLB", "ILB", "OLB", "LB"))
        if(nrow(lbs) > 0 && next_slot <= 11) {
          for(i in 1:min(11 - next_slot + 1, nrow(lbs))) {
            player_slots[next_slot + i - 1] <- lbs$gsis_id[i]
          }
        }
        
        # Find next empty slot again
        next_empty <- which(is.na(player_slots))
        if(length(next_empty) > 0) {
          next_slot <- min(next_empty)
          
          # Secondary last
          secondary <- players_info %>% 
            filter(depth_chart_position %in% c("S", "FS", "SS", "SAF", "CB", "DB"))
          if(nrow(secondary) > 0 && next_slot <= 11) {
            for(i in 1:min(11 - next_slot + 1, nrow(secondary))) {
              player_slots[next_slot + i - 1] <- secondary$gsis_id[i]
            }
          }
        }
      }
    }
    
    return(player_slots)
  }
  
  # Apply position assignments to third downs
  message("  Processing third downs...")
  for(i in 1:nrow(third_downs)) {
    if(i %% 1000 == 0) message("    Processing row ", i, " of ", nrow(third_downs))
    
    # Process offense
    if(!is.na(third_downs$offense_players[i])) {
      player_slots <- assign_players_to_positions(third_downs$offense_players[i], roster_data, TRUE)
      for(j in 1:11) {
        third_downs[[paste0("offense_player_", j)]][i] <- player_slots[j]
      }
    }
    
    # Process defense
    if(!is.na(third_downs$defense_players[i])) {
      player_slots <- assign_players_to_positions(third_downs$defense_players[i], roster_data, FALSE)
      for(j in 1:11) {
        third_downs[[paste0("defense_player_", j)]][i] <- player_slots[j]
      }
    }
  }
  
  # Apply position assignments to fourth downs
  message("  Processing fourth downs...")
  for(i in 1:nrow(fourth_downs)) {
    if(i %% 1000 == 0) message("    Processing row ", i, " of ", nrow(fourth_downs))
    
    # Process offense
    if(!is.na(fourth_downs$offense_players[i])) {
      player_slots <- assign_players_to_positions(fourth_downs$offense_players[i], roster_data, TRUE)
      for(j in 1:11) {
        fourth_downs[[paste0("offense_player_", j)]][i] <- player_slots[j]
      }
    }
    
    # Process defense
    if(!is.na(fourth_downs$defense_players[i])) {
      player_slots <- assign_players_to_positions(fourth_downs$defense_players[i], roster_data, FALSE)
      for(j in 1:11) {
        fourth_downs[[paste0("defense_player_", j)]][i] <- player_slots[j]
      }
    }
  }
  
  message("Step 3.7/7: Counting original and assigned players...")
  
  # For third downs
  message("  Processing third downs...")
  third_downs <- third_downs %>%
    mutate(
      original_offense_count = ifelse(is.na(offense_players) | offense_players == "", 
                                    0, 
                                    lengths(strsplit(offense_players, ";"))),
      original_defense_count = ifelse(is.na(defense_players) | defense_players == "", 
                                    0, 
                                    lengths(strsplit(defense_players, ";")))
    ) %>%
    rowwise() %>%
    mutate(
      assigned_offense_count = sum(!is.na(c_across(starts_with("offense_player_")))),
      assigned_defense_count = sum(!is.na(c_across(starts_with("defense_player_"))))
    ) %>%
    ungroup()
  
  # For fourth downs
  message("  Processing fourth downs...")
  fourth_downs <- fourth_downs %>%
    mutate(
      original_offense_count = ifelse(is.na(offense_players) | offense_players == "", 
                                    0, 
                                    lengths(strsplit(offense_players, ";"))),
      original_defense_count = ifelse(is.na(defense_players) | defense_players == "", 
                                    0, 
                                    lengths(strsplit(defense_players, ";")))
    ) %>%
    rowwise() %>%
    mutate(
      assigned_offense_count = sum(!is.na(c_across(starts_with("offense_player_")))),
      assigned_defense_count = sum(!is.na(c_across(starts_with("defense_player_"))))
    ) %>%
    ungroup()

  message("Step 4/7: Adding player information...")
  for(i in 1:11) {
    cat("\rProcessing player", i, "of 11")
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("offense_player_", i), 
                                  paste0("offense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("offense_player_", i), 
                                   paste0("offense_player_", i))
    
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("defense_player_", i), 
                                  paste0("defense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("defense_player_", i), 
                                   paste0("defense_player_", i))
  }

  message("Step 5/7: Processing team injury data...")
  third_downs <- third_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  fourth_downs <- fourth_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  third_downs <- third_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  fourth_downs <- fourth_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  third_downs <- third_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  fourth_downs <- fourth_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  all_teams_weeks <- bind_rows(
    select(third_downs, defteam, week),
    select(fourth_downs, defteam, week)
  ) %>% distinct()

  prev_stats <- map2_dfr(
    all_teams_weeks$defteam,
    all_teams_weeks$week,
    ~{
      prev_game <- play_year %>%
        filter((home_team == .x | away_team == .x) & week < .y) %>%
        arrange(desc(week)) %>%
        slice(1) %>%
        pull(week)
      
      if(length(prev_game) == 0) return(tibble(
        defteam = .x,
        week = .y,
        attempts = 0,
        successes = 0,
        failures = 0
      ))
      
      play_year %>%
        filter(week == prev_game,
               (home_team == .x | away_team == .x),
               down == 4,
               !play_type %in% c("punt", "field_goal")) %>%
        summarize(
          defteam = .x,
          week = .y,
          attempts = n(),
          successes = sum(first_down == 1, na.rm = TRUE),
          failures = sum(first_down == 0, na.rm = TRUE)
        )
    })

  message("Step 5.9/7: Debugging data structures section...")


# Add fixed down values since all third downs have down=3
third_downs$down <- 3
fourth_downs$down <- 4


  
message("Step 6 Diagnostics: Checking key data structures and columns...")



  model_3rd <- third_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  model_4th <- fourth_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
    left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  
  # Add before returning the models in process_year:
  model_3rd <- model_3rd %>%
    left_join(coach_records, by = c("posteam_coach" = "coach")) %>%
    left_join(current_team_records, by = c("posteam_coach" = "coach", "posteam")) %>%
    left_join(game_tenure %>% 
                select(game_id, posteam, coach, tenure_days), 
              by = c("game_id", "posteam", "posteam_coach" = "coach")) %>%
    left_join(formation_tendencies %>% 
                filter(down == 3), 
              by = c("posteam", "down")) %>%
    left_join(defense_tendencies %>% 
                filter(down == 3), 
              by = c("defteam", "down")) %>%
    left_join(coverage_tendencies %>% 
                filter(down == 3), 
              by = c("defteam", "defense_coverage_type"))

  model_4th <- model_4th %>%
    left_join(coach_records, by = c("posteam_coach" = "coach")) %>%
    left_join(current_team_records, by = c("posteam_coach" = "coach", "posteam")) %>%
    left_join(game_tenure %>% 
                select(game_id, posteam, coach, tenure_days), 
              by = c("game_id", "posteam", "posteam_coach" = "coach")) %>%
    left_join(formation_tendencies %>% 
                filter(down == 4), 
              by = c("posteam", "down")) %>%
    left_join(defense_tendencies %>% 
                filter(down == 4), 
              by = c("defteam", "down")) %>%
    left_join(coverage_tendencies %>% 
                filter(down == 4), 
              by = c("defteam", "defense_coverage_type"))
  
  # Add after final model joins in Step 6:
message("Diagnosing final model duplicates...")
model_4th_dupes <- model_4th %>%
  group_by(play_id, game_id) %>%
  filter(n() > 1)

print("Number of duplicate rows in final fourth down model:")
print(nrow(model_4th_dupes))
if(nrow(model_4th_dupes) > 0) {
  print("Sample of duplicates with key columns:")
  print(model_4th_dupes %>% 
        select(play_id, game_id, posteam, defteam, offense_formation, defense_coverage_type) %>%
        head())
}
  
  
  message("Step 7/7: Adding specialists...")
  specialists <- get_specialists(year)

  specialists$kickers <- specialists$kickers %>%
    rename(
      kicker_id = kicker_player_id,
      kicker_name = kicker_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, kicker_id, kicker_name)

  specialists$punters <- specialists$punters %>%
    rename(
      punter_id = punter_player_id,
      punter_name = punter_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, punter_id, punter_name)

  model_4th <- model_4th %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  model_3rd <- model_3rd %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  return(list(
    third_downs = model_3rd,
    fourth_downs = model_4th
  ))
}
```

```{r}
# Main processing
years <- 2016:2023
all_data <- list()

for(year in years) {
  all_data[[as.character(year)]] <- process_year(year)
}

combined_third_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$third_downs))
)

combined_fourth_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$fourth_downs))
)
```

```{r}
# Save results
write.csv(combined_third_downs, "ctd.csv", row.names = FALSE)
write.csv(combined_fourth_downs, "cfd.csv", row.names = FALSE)
```


next is coaches

```{r}
library(readr)
#combined_third_downs <- read.csv("ctd.csv")
#combined_fourth_downs <- read.csv("cfd.csv")
```


# Explinations and proofs

offense_player_1_days_since_injury       # Days since last injury for player 

offense_player_1_current_status         # Most recent game status (Questionable/Doubtful/Out/NA)
offense_player_1_current_practice       # Most recent practice status (Limited/DNP/Full/NA)

these look 1 week back

offense_player_1_week_questionable      # Number of times listed as "Questionable" on injury report
offense_player_1_week_doubtful          # Number of times listed as "Doubtful" on injury report  
offense_player_1_week_out               # Number of times listed as "Out" on injury report
offense_player_1_week_limited_practice  # Number of times had "Limited Participation in Practice" 
offense_player_1_week_dnp               # Number of times "Did Not Participate in Practice"




# Filtering for only 4th down attempts that are a rush_attempt = 1 or pass_attempt = 1
```{r}
attempts_4th <- combined_fourth_downs %>%
  filter(rush_attempt == 1 | pass_attempt == 1)


```

```{r}
# Define defensive positions
defensive_positions <- c("K", "P", "LS", "DB", "CB", "S", "FS", "SS", "LB", "ILB", "OLB", "MLB", "DL", "DE", "DT", "NT", "EDGE")

# Function to check if any offensive player has a defensive position
check_defensive_positions <- function(row) {
  # Get all offensive player positions
  positions <- c()
  for(i in 1:11) {
    pos_col <- paste0("offense_player_", i, "_position")
    if(pos_col %in% names(row)) {
      positions <- c(positions, row[[pos_col]])
    }
  }
  
  # Return TRUE if no defensive positions found
  !any(positions %in% defensive_positions, na.rm = TRUE)
}

# Apply the filter
clean_attempts_4th <- attempts_4th %>%
  filter(mapply(check_defensive_positions, split(., 1:nrow(.))))

# Print the number of rows removed
cat("Original rows:", nrow(attempts_4th), "\n")
cat("Rows after filtering:", nrow(clean_attempts_4th), "\n")
cat("Rows removed:", nrow(attempts_4th) - nrow(clean_attempts_4th), "\n")

# Optional: Examine some examples of removed plays
removed_plays <- attempts_4th %>%
  filter(!mapply(check_defensive_positions, split(., 1:nrow(.)))) %>%
  select(play_id, game_id, desc, matches("offense_player_\\d+_position"))

# Show first few removed plays
head(removed_plays)
```


