---
title: "pbp_data_creation"
output: html_document
date: "2024-12-04"
---



```{r}
# Load required libraries
library(dplyr)
library(nflverse)
library(tidyr)
library(purrr)
library(zoo)
library(car)
library(progress)


special_teams_order <- c("K", "P", "LS")

# Validation function
validate_data <- function(df) {
  dupes <- duplicated(names(df))
  if(any(dupes)) {
    warning("Duplicate columns found: ", paste(names(df)[dupes], collapse=", "))
  }
  
  req_cols <- c("play_id", "game_id", "posteam")
  missing <- req_cols[!req_cols %in% names(df)]
  if(length(missing) > 0) {
    stop("Missing required columns: ", paste(missing, collapse=", "))
  }
  return(df)
}


# Get specialists function
get_specialists <- function(year) {
  pbp <- bind_rows(
    load_pbp(year),
    load_pbp(year - 1)
  ) %>%
  arrange(game_date, desc(play_id))
  
  kickers <- pbp %>%
    filter(field_goal_attempt == 1 | extra_point_attempt == 1) %>%
    select(team = posteam, week, season, kicker_player_id, kicker_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  punters <- pbp %>%
    filter(punt_attempt == 1) %>%
    select(team = posteam, week, season, punter_player_id, punter_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  list(kickers = kickers, punters = punters)
}
add_player_info <- function(df, players_data, player_col, prefix) {
    if (!player_col %in% names(df)) {
        return(df)
    }
    
    new_cols <- c("full_name", "birth_date", "rookie_year", "draft_number", "jersey_number", "pff_id", 
                "weight", "height", "position", "depth_chart_position")
    
    for(col in new_cols) {
        new_col_name <- paste0(prefix, "_", col)
        df[[new_col_name]] <- NA
    }
    
    df[[paste0(prefix, "_days_since_injury")]] <- NA
    df[[paste0(prefix, "_last_injury_type")]] <- NA
    df[[paste0(prefix, "_injury_count_1yr")]] <- NA
    
    df[[paste0(prefix, "_week_questionable")]] <- 0
    df[[paste0(prefix, "_week_doubtful")]] <- 0
    df[[paste0(prefix, "_week_out")]] <- 0
    df[[paste0(prefix, "_week_limited_practice")]] <- 0
    df[[paste0(prefix, "_week_dnp")]] <- 0
    df[[paste0(prefix, "_current_status")]] <- NA
    df[[paste0(prefix, "_current_practice")]] <- NA
    
    valid_players <- unique(df[[player_col]][!is.na(df[[player_col]])])
    
    if(length(valid_players) > 0) {
        players_info <- players_data %>%
            filter(gsis_id %in% valid_players) %>%
            group_by(gsis_id) %>%
            slice_tail(n = 1) %>%
            ungroup() %>%
            select(gsis_id, all_of(new_cols))
        
        for(i in 1:nrow(df)) {
            if(!is.na(df[[player_col]][i])) {
                player_id <- df[[player_col]][i]
                current_date <- as.Date(df$game_date[i])
                player_data <- players_info %>% filter(gsis_id == player_id)
                
                if(nrow(player_data) == 1) {
                    for(col in new_cols) {
                        if(!is.null(player_data[[col]])) {
                            df[i, paste0(prefix, "_", col)] <- player_data[[col]][1]
                        }
                    }
                }
                
                player_injuries <- injuries_data %>%
                    filter(gsis_id == player_id,
                           injury_date < current_date,
                           !is.na(injury_date)) %>%
                    arrange(desc(injury_date))
                
                if(nrow(player_injuries) > 0) {
                    df[i, paste0(prefix, "_days_since_injury")] <- 
                        as.numeric(difftime(current_date, player_injuries$injury_date[1], units = "days"))
                    
                    df[i, paste0(prefix, "_last_injury_type")] <- 
                        player_injuries$practice_primary_injury[1]
                    
                    one_year_ago <- current_date - 365
                    df[i, paste0(prefix, "_injury_count_1yr")] <- 
                        sum(player_injuries$injury_date >= one_year_ago)
                    
                    week_ago <- current_date - 7
                    recent_injuries <- injuries_data %>%
                        filter(gsis_id == player_id,
                               injury_date >= week_ago,
                               injury_date < current_date)
                    
                    if(nrow(recent_injuries) > 0) {
                        df[i, paste0(prefix, "_week_questionable")] <- 
                            sum(recent_injuries$report_status == "Questionable", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_doubtful")] <- 
                            sum(recent_injuries$report_status == "Doubtful", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_out")] <- 
                            sum(recent_injuries$report_status == "Out", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_limited_practice")] <- 
                            sum(recent_injuries$practice_status == "Limited Participation in Practice", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_dnp")] <- 
                            sum(recent_injuries$practice_status == "Did Not Participate In Practice", na.rm = TRUE)
                        
                        df[i, paste0(prefix, "_current_status")] <- recent_injuries$report_status[1]
                        df[i, paste0(prefix, "_current_practice")] <- recent_injuries$practice_status[1]
                    }
                }
            }
        }
    }
    return(df)
}
```


```{r}
process_year <- function(year) {
 message(paste("\nProcessing year:", year))
  
  message("Step 1/7: Loading data...")
  # Add global assignment
  injuries_data <<- load_injuries((year-2):year) %>%
    mutate(injury_date = as.Date(date_modified))

  top_injury_types <- injuries_data %>%
    filter(!is.na(practice_primary_injury)) %>%
    count(practice_primary_injury, sort = TRUE) %>%
    slice_head(n = 9) %>%
    pull(practice_primary_injury)

  get_team_injury_counts <- function(team, game_date) {
    game_date <- as.Date(game_date)
    
    team_injuries <- injuries_data %>%
      filter(team == !!team) %>%
      mutate(
        injury_date = as.Date(date_modified),
        standard_injury = case_when(
          practice_primary_injury %in% top_injury_types ~ practice_primary_injury,
          TRUE ~ "Other"
        )
      )
    
    year_ago <- game_date - 365
    year_injuries <- team_injuries %>%
      filter(injury_date >= year_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_year_",
        values_from = n,
        values_fill = 0
      )
    
    week_ago <- game_date - 7
    week_injuries <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_week_",
        values_from = n,
        values_fill = 0
      )
      
    week_status <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      summarise(
        week_questionable = sum(report_status == "Questionable", na.rm = TRUE),
        week_doubtful = sum(report_status == "Doubtful", na.rm = TRUE),
        week_out = sum(report_status == "Out", na.rm = TRUE),
        week_limited_practice = sum(practice_status == "Limited Participation in Practice", na.rm = TRUE),
        week_dnp = sum(practice_status == "Did Not Participate In Practice", na.rm = TRUE)
      )
    
    bind_cols(year_injuries, week_injuries, week_status)
  }

  play_year <- load_pbp(year)
  participation_data <- load_participation(year) %>%
    select(
      nflverse_game_id,
      play_id,
      offense_formation,
      offense_personnel,
      defense_personnel,
      defenders_in_box,
      number_of_pass_rushers,
      offense_players,
      defense_players,
      defense_man_zone_type,
      defense_coverage_type
    )
    
  roster_data <- load_rosters(year)
    
  message("Step 2/7: Processing weekly stats...")
  weekly_run_stats <- play_year %>%
    filter(play_type %in% c("pass", "run"), down %in% 1:3) %>%
    group_by(posteam, week, down) %>%
    summarize(
      weekly_run_plays = sum(play_type == "run", na.rm = TRUE),
      weekly_total_plays = n(),
      .groups = 'drop'
    ) %>%
    arrange(posteam, week, down) %>%
    group_by(posteam, down) %>%
    mutate(
      cum_run_plays = cumsum(weekly_run_plays),
      cum_total_plays = cumsum(weekly_total_plays)
    ) %>%
    ungroup() %>%
    pivot_wider(
      id_cols = c(posteam, week),
      names_from = down,
      names_glue = "down{down}_{.value}",
      values_from = c(cum_run_plays, cum_total_plays)
    ) %>%
    arrange(posteam, week)

  run_pcts <- weekly_run_stats %>%
    mutate(
      down1_pct = round(100 * down1_cum_run_plays / down1_cum_total_plays, 1),
      down2_pct = round(100 * down2_cum_run_plays / down2_cum_total_plays, 1),
      down3_pct = round(100 * down3_cum_run_plays / down3_cum_total_plays, 1)
    ) %>%
    select(posteam, week, down1_pct, down2_pct, down3_pct)

  prep_time_df <- play_year %>%
    select(game_id, game_date, posteam = home_team) %>%
    distinct() %>%
    bind_rows(
      play_year %>%
        select(game_id, game_date, posteam = away_team) %>%
        distinct()
    ) %>%
    arrange(posteam, game_date) %>%
    group_by(posteam) %>%
    mutate(
      last_game_date = lag(game_date),
      prep_days = as.numeric(difftime(game_date, last_game_date, units = "days"))
    ) %>%
    ungroup()

  message("Step 3/7: Processing third and fourth downs...")
  third_downs <- play_year %>%
    filter(down == 3) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      season = year
    ) %>%
    filter(!field_goal_attempt, !punt_attempt, play_type %in% c("pass", "run")) %>%
    mutate(converted = first_down) %>%
    select(-first_down) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id")) %>%
    separate(offense_players, 
             into = paste0("offense_player_", 1:11), 
             sep = ";", 
             fill = "right") %>%
    separate(defense_players, 
             into = paste0("defense_player_", 1:11), 
             sep = ";", 
             fill = "right")

  fourth_downs <- play_year %>%
    filter(down == 4) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      attempt = if_else(play_type %in% c("punt", "field_goal"), 0, 1),
      season = year
    ) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id")) %>%
    separate(offense_players, 
             into = paste0("offense_player_", 1:11), 
             sep = ";", 
             fill = "right") %>%
    separate(defense_players, 
             into = paste0("defense_player_", 1:11), 
             sep = ";", 
             fill = "right")

  max_week <- if(year >= 2021) 18 else 17

  third_downs <- third_downs %>%
    filter(week <= max_week)

  fourth_downs <- fourth_downs %>%
    filter(week <= max_week)
    
  message("Step 4/7: Adding player information...")
  for(i in 1:11) {
    cat("\rProcessing player", i, "of 11")
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("offense_player_", i), 
                                  paste0("offense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("offense_player_", i), 
                                   paste0("offense_player_", i))
    
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("defense_player_", i), 
                                  paste0("defense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("defense_player_", i), 
                                   paste0("defense_player_", i))
  }

  message("Step 5/7: Processing team injury data...")
  third_downs <- third_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  fourth_downs <- fourth_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  third_downs <- third_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  fourth_downs <- fourth_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  third_downs <- third_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  fourth_downs <- fourth_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  all_teams_weeks <- bind_rows(
    select(third_downs, defteam, week),
    select(fourth_downs, defteam, week)
  ) %>% distinct()

  prev_stats <- map2_dfr(
    all_teams_weeks$defteam,
    all_teams_weeks$week,
    ~{
      prev_game <- play_year %>%
        filter((home_team == .x | away_team == .x) & week < .y) %>%
        arrange(desc(week)) %>%
        slice(1) %>%
        pull(week)
      
      if(length(prev_game) == 0) return(tibble(
        defteam = .x,
        week = .y,
        attempts = 0,
        successes = 0,
        failures = 0
      ))
      
      play_year %>%
        filter(week == prev_game,
               (home_team == .x | away_team == .x),
               down == 4,
               !play_type %in% c("punt", "field_goal")) %>%
        summarize(
          defteam = .x,
          week = .y,
          attempts = n(),
          successes = sum(first_down == 1, na.rm = TRUE),
          failures = sum(first_down == 0, na.rm = TRUE)
        )
    })

  message("Step 6/7: Processing model data...")
  model_3rd <- third_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  model_4th <- fourth_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
    left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  message("Step 7/7: Adding specialists...")
  specialists <- get_specialists(year)

  specialists$kickers <- specialists$kickers %>%
    rename(
      kicker_id = kicker_player_id,
      kicker_name = kicker_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, kicker_id, kicker_name)

  specialists$punters <- specialists$punters %>%
    rename(
      punter_id = punter_player_id,
      punter_name = punter_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, punter_id, punter_name)

  model_4th <- model_4th %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  model_3rd <- model_3rd %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  return(list(
    third_downs = model_3rd,
    fourth_downs = model_4th
  ))
}
```

```{r}
# Main processing
years <- 2016:2023
all_data <- list()

for(year in years) {
  all_data[[as.character(year)]] <- process_year(year)
}

combined_third_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$third_downs))
)

combined_fourth_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$fourth_downs))
)
```

```{r}
# Save results
#write.csv(combined_third_downs, "ctd.csv", row.names = FALSE)
#write.csv(combined_fourth_downs, "cfd.csv", row.names = FALSE)
```


next is coaches

```{r}
library(readr)
#combined_third_downs <- read.csv("ctd.csv")
#combined_fourth_downs <- read.csv("cfd.csv")
```


# Explinations and proofs

offense_player_1_days_since_injury       # Days since last injury for player 

offense_player_1_current_status         # Most recent game status (Questionable/Doubtful/Out/NA)
offense_player_1_current_practice       # Most recent practice status (Limited/DNP/Full/NA)

these look 1 week back

offense_player_1_week_questionable      # Number of times listed as "Questionable" on injury report
offense_player_1_week_doubtful          # Number of times listed as "Doubtful" on injury report  
offense_player_1_week_out               # Number of times listed as "Out" on injury report
offense_player_1_week_limited_practice  # Number of times had "Limited Participation in Practice" 
offense_player_1_week_dnp               # Number of times "Did Not Participate in Practice"




# Filtering for only 4th down attempts that are a rush_attempt = 1 or pass_attempt = 1
```{r}
attempts_4th <- combined_fourth_downs %>%
  filter(rush_attempt == 1 | pass_attempt == 1)


```

```{r}
# Define defensive positions
defensive_positions <- c("K", "P", "LS", "DB", "CB", "S", "FS", "SS", "LB", "ILB", "OLB", "MLB", "DL", "DE", "DT", "NT", "EDGE")

# Function to check if any offensive player has a defensive position
check_defensive_positions <- function(row) {
  # Get all offensive player positions
  positions <- c()
  for(i in 1:11) {
    pos_col <- paste0("offense_player_", i, "_position")
    if(pos_col %in% names(row)) {
      positions <- c(positions, row[[pos_col]])
    }
  }
  
  # Return TRUE if no defensive positions found
  !any(positions %in% defensive_positions, na.rm = TRUE)
}

# Apply the filter
clean_attempts_4th <- attempts_4th %>%
  filter(mapply(check_defensive_positions, split(., 1:nrow(.))))

# Print the number of rows removed
cat("Original rows:", nrow(attempts_4th), "\n")
cat("Rows after filtering:", nrow(clean_attempts_4th), "\n")
cat("Rows removed:", nrow(attempts_4th) - nrow(clean_attempts_4th), "\n")

# Optional: Examine some examples of removed plays
removed_plays <- attempts_4th %>%
  filter(!mapply(check_defensive_positions, split(., 1:nrow(.)))) %>%
  select(play_id, game_id, desc, matches("offense_player_\\d+_position"))

# Show first few removed plays
head(removed_plays)
```


