---
title: "pbp_data_creation"
output: html_document
date: "2024-12-04"
---



```{r}
# Load required libraries
library(dplyr)
library(nflverse)
library(tidyr)
library(purrr)
library(zoo)
library(car)
library(progress)

# Validation function
validate_data <- function(df) {
  dupes <- duplicated(names(df))
  if(any(dupes)) {
    warning("Duplicate columns found: ", paste(names(df)[dupes], collapse=", "))
  }
  
  req_cols <- c("play_id", "game_id", "posteam")
  missing <- req_cols[!req_cols %in% names(df)]
  if(length(missing) > 0) {
    stop("Missing required columns: ", paste(missing, collapse=", "))
  }
  return(df)
}

# Get specialists function
get_specialists <- function(year) {
  pbp <- bind_rows(
    load_pbp(year),
    load_pbp(year - 1)
  ) %>%
  arrange(game_date, desc(play_id))
  
  kickers <- pbp %>%
    filter(field_goal_attempt == 1 | extra_point_attempt == 1) %>%
    select(team = posteam, week, season, kicker_player_id, kicker_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  punters <- pbp %>%
    filter(punt_attempt == 1) %>%
    select(team = posteam, week, season, punter_player_id, punter_player_name) %>%
    distinct(team, week, season, .keep_all = TRUE)
    
  list(kickers = kickers, punters = punters)
}
add_player_info <- function(df, players_data, player_col, prefix) {
    if (!player_col %in% names(df)) {
        return(df)
    }
    
    new_cols <- c("full_name", "birth_date", "rookie_year", "draft_number", "jersey_number", "pff_id", 
                "weight", "height", "position", "depth_chart_position")
    
    for(col in new_cols) {
        new_col_name <- paste0(prefix, "_", col)
        df[[new_col_name]] <- NA
    }
    
    df[[paste0(prefix, "_days_since_injury")]] <- NA
    df[[paste0(prefix, "_last_injury_type")]] <- NA
    df[[paste0(prefix, "_injury_count_1yr")]] <- NA
    
    df[[paste0(prefix, "_week_questionable")]] <- 0
    df[[paste0(prefix, "_week_doubtful")]] <- 0
    df[[paste0(prefix, "_week_out")]] <- 0
    df[[paste0(prefix, "_week_limited_practice")]] <- 0
    df[[paste0(prefix, "_week_dnp")]] <- 0
    df[[paste0(prefix, "_current_status")]] <- NA
    df[[paste0(prefix, "_current_practice")]] <- NA
    
    valid_players <- unique(df[[player_col]][!is.na(df[[player_col]])])
    
    if(length(valid_players) > 0) {
        players_info <- players_data %>%
            filter(gsis_id %in% valid_players) %>%
            group_by(gsis_id) %>%
            slice_tail(n = 1) %>%
            ungroup() %>%
            select(gsis_id, all_of(new_cols))
        
        for(i in 1:nrow(df)) {
            if(!is.na(df[[player_col]][i])) {
                player_id <- df[[player_col]][i]
                current_date <- as.Date(df$game_date[i])
                player_data <- players_info %>% filter(gsis_id == player_id)
                
                if(nrow(player_data) == 1) {
                    for(col in new_cols) {
                        if(!is.null(player_data[[col]])) {
                            df[i, paste0(prefix, "_", col)] <- player_data[[col]][1]
                        }
                    }
                }
                
                player_injuries <- injuries_data %>%
                    filter(gsis_id == player_id,
                           injury_date < current_date,
                           !is.na(injury_date)) %>%
                    arrange(desc(injury_date))
                
                if(nrow(player_injuries) > 0) {
                    df[i, paste0(prefix, "_days_since_injury")] <- 
                        as.numeric(difftime(current_date, player_injuries$injury_date[1], units = "days"))
                    
                    df[i, paste0(prefix, "_last_injury_type")] <- 
                        player_injuries$practice_primary_injury[1]
                    
                    one_year_ago <- current_date - 365
                    df[i, paste0(prefix, "_injury_count_1yr")] <- 
                        sum(player_injuries$injury_date >= one_year_ago)
                    
                    week_ago <- current_date - 7
                    recent_injuries <- injuries_data %>%
                        filter(gsis_id == player_id,
                               injury_date >= week_ago,
                               injury_date < current_date)
                    
                    if(nrow(recent_injuries) > 0) {
                        df[i, paste0(prefix, "_week_questionable")] <- 
                            sum(recent_injuries$report_status == "Questionable", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_doubtful")] <- 
                            sum(recent_injuries$report_status == "Doubtful", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_out")] <- 
                            sum(recent_injuries$report_status == "Out", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_limited_practice")] <- 
                            sum(recent_injuries$practice_status == "Limited Participation in Practice", na.rm = TRUE)
                        df[i, paste0(prefix, "_week_dnp")] <- 
                            sum(recent_injuries$practice_status == "Did Not Participate In Practice", na.rm = TRUE)
                        
                        df[i, paste0(prefix, "_current_status")] <- recent_injuries$report_status[1]
                        df[i, paste0(prefix, "_current_practice")] <- recent_injuries$practice_status[1]
                    }
                }
            }
        }
    }
    return(df)
}
```


```{r}
process_year <- function(year) {
 message(paste("\nProcessing year:", year))
  
  message("Step 1/7: Loading data...")
  # Add global assignment
  injuries_data <<- load_injuries((year-2):year) %>%
    mutate(injury_date = as.Date(date_modified))

  top_injury_types <- injuries_data %>%
    filter(!is.na(practice_primary_injury)) %>%
    count(practice_primary_injury, sort = TRUE) %>%
    slice_head(n = 9) %>%
    pull(practice_primary_injury)

  get_team_injury_counts <- function(team, game_date) {
    game_date <- as.Date(game_date)
    
    team_injuries <- injuries_data %>%
      filter(team == !!team) %>%
      mutate(
        injury_date = as.Date(date_modified),
        standard_injury = case_when(
          practice_primary_injury %in% top_injury_types ~ practice_primary_injury,
          TRUE ~ "Other"
        )
      )
    
    year_ago <- game_date - 365
    year_injuries <- team_injuries %>%
      filter(injury_date >= year_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_year_",
        values_from = n,
        values_fill = 0
      )
    
    week_ago <- game_date - 7
    week_injuries <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      count(standard_injury) %>%
      pivot_wider(
        names_from = standard_injury,
        names_prefix = "injury_count_week_",
        values_from = n,
        values_fill = 0
      )
      
    week_status <- team_injuries %>%
      filter(injury_date >= week_ago,
             injury_date < game_date) %>%
      summarise(
        week_questionable = sum(report_status == "Questionable", na.rm = TRUE),
        week_doubtful = sum(report_status == "Doubtful", na.rm = TRUE),
        week_out = sum(report_status == "Out", na.rm = TRUE),
        week_limited_practice = sum(practice_status == "Limited Participation in Practice", na.rm = TRUE),
        week_dnp = sum(practice_status == "Did Not Participate In Practice", na.rm = TRUE)
      )
    
    bind_cols(year_injuries, week_injuries, week_status)
  }

  play_year <- load_pbp(year)
  participation_data <- load_participation(year) %>%
    select(
      nflverse_game_id,
      play_id,
      offense_formation,
      offense_personnel,
      defense_personnel,
      defenders_in_box,
      number_of_pass_rushers,
      offense_players,
      defense_players,
      defense_man_zone_type,
      defense_coverage_type
    )
    
  roster_data <- load_rosters(year)
    
  message("Step 2/7: Processing weekly stats...")
  weekly_run_stats <- play_year %>%
    filter(play_type %in% c("pass", "run"), down %in% 1:3) %>%
    group_by(posteam, week, down) %>%
    summarize(
      weekly_run_plays = sum(play_type == "run", na.rm = TRUE),
      weekly_total_plays = n(),
      .groups = 'drop'
    ) %>%
    arrange(posteam, week, down) %>%
    group_by(posteam, down) %>%
    mutate(
      cum_run_plays = cumsum(weekly_run_plays),
      cum_total_plays = cumsum(weekly_total_plays)
    ) %>%
    ungroup() %>%
    pivot_wider(
      id_cols = c(posteam, week),
      names_from = down,
      names_glue = "down{down}_{.value}",
      values_from = c(cum_run_plays, cum_total_plays)
    ) %>%
    arrange(posteam, week)

  run_pcts <- weekly_run_stats %>%
    mutate(
      down1_pct = round(100 * down1_cum_run_plays / down1_cum_total_plays, 1),
      down2_pct = round(100 * down2_cum_run_plays / down2_cum_total_plays, 1),
      down3_pct = round(100 * down3_cum_run_plays / down3_cum_total_plays, 1)
    ) %>%
    select(posteam, week, down1_pct, down2_pct, down3_pct)

  prep_time_df <- play_year %>%
    select(game_id, game_date, posteam = home_team) %>%
    distinct() %>%
    bind_rows(
      play_year %>%
        select(game_id, game_date, posteam = away_team) %>%
        distinct()
    ) %>%
    arrange(posteam, game_date) %>%
    group_by(posteam) %>%
    mutate(
      last_game_date = lag(game_date),
      prep_days = as.numeric(difftime(game_date, last_game_date, units = "days"))
    ) %>%
    ungroup()

  # Add after run_pcts but before prep_time_df in Step 2/7:
  message("Step 2.5/7: Processing coach history...")
  
  # Get coach records across all games
  coach_records <- play_year %>%
    select(game_id, game_date, home_team, away_team, home_coach, away_coach,
           home_score, away_score) %>%
    distinct() %>%
    # Get home coach records
    mutate(
      home_win = home_score > away_score,
      home_loss = home_score < away_score
    ) %>%
    group_by(home_coach) %>%
    summarize(
      overall_wins = sum(home_win, na.rm = TRUE),
      overall_losses = sum(home_loss, na.rm = TRUE),
      overall_win_pct = round(sum(home_win, na.rm = TRUE) / n(), 3)
    ) %>%
    # Combine with away coach records
    bind_rows(
      play_year %>%
      select(game_id, away_coach, home_score, away_score) %>%
      distinct() %>%
      group_by(away_coach) %>%
      summarize(
        overall_wins = sum(away_score > home_score, na.rm = TRUE),
        overall_losses = sum(away_score < home_score, na.rm = TRUE),
        overall_win_pct = round(sum(away_score > home_score, na.rm = TRUE) / n(), 3)
      )
    ) %>%
    group_by(home_coach) %>%
    summarize(
      overall_wins = sum(overall_wins),
      overall_losses = sum(overall_losses),
      overall_win_pct = round(mean(overall_win_pct), 3)
    ) %>%
    rename(coach = home_coach)

  # Get coach records with current team only
  current_team_records <- play_year %>%
    select(game_id, game_date, posteam, home_team, away_team, home_coach, away_coach,
           posteam_score, defteam_score) %>%
    distinct() %>%
    mutate(
      coach = ifelse(posteam == home_team, home_coach, away_coach)
    ) %>%
    group_by(coach, posteam) %>%
    summarize(
      team_wins = sum(posteam_score > defteam_score, na.rm = TRUE),
      team_losses = sum(posteam_score < defteam_score, na.rm = TRUE),
      team_win_pct = round(mean(posteam_score > defteam_score, na.rm = TRUE), 3),
      .groups = 'drop'
    )
  
  # Add after coach records calculation in Step 2.5:
message("Diagnosing coach records duplicates...")
coach_dupes <- coach_records %>%
  group_by(coach) %>%
  filter(n() > 1)

print("Number of duplicate coach record rows:")
print(nrow(coach_dupes))
if(nrow(coach_dupes) > 0) {
  print("Sample of duplicates:")
  print(head(coach_dupes))
}
  
 message("Step 2.6/7: Processing coach tenure...")
# Calculate coach tenure - modify this section
coach_tenure <- play_year %>%
    select(game_date, home_team, away_team, home_coach, away_coach) %>%
    # Get unique games first
    distinct() %>%
    pivot_longer(
      cols = c(home_coach, away_coach),
      names_to = "role",
      values_to = "coach"
    ) %>%
    mutate(team = if_else(role == "home_coach", home_team, away_team)) %>%
    # Get first appearance for each coach-team combo
    group_by(coach, team) %>%
    summarize(
      first_appearance = min(game_date),
      .groups = 'drop'
    )

# Calculate tenure for each game - modify this section
game_tenure <- play_year %>%
    select(game_id, game_date, posteam, home_team, away_team, home_coach, away_coach) %>%
    # Get unique game-team combinations
    distinct() %>%
    mutate(
      coach = ifelse(posteam == home_team, home_coach, away_coach)
    ) %>%
    left_join(coach_tenure, by = c("coach", "posteam" = "team")) %>%
    mutate(
      tenure_days = as.numeric(as.Date(game_date) - as.Date(first_appearance))
    )
  
  # Add after game tenure calculation in Step 2.6:
message("Diagnosing game tenure duplicates...")
tenure_dupes <- game_tenure %>%
  group_by(game_id, posteam, coach) %>%
  filter(n() > 1)

print("Number of duplicate tenure rows:")
print(nrow(tenure_dupes))
if(nrow(tenure_dupes) > 0) {
  print("Sample of duplicates:")
  print(head(tenure_dupes))
}

# Add after the game tenure diagnostic check:
if(nrow(tenure_dupes) > 0) {
    print("Detailed look at tenure duplicates:")
    print(
        tenure_dupes %>%
        arrange(game_id, posteam, coach) %>%
        select(game_id, game_date, posteam, coach, first_appearance, tenure_days) %>%
        head(10)
    )
    
    # Count frequency of duplications
    print("Number of duplicates per game/team/coach combination:")
    print(
        tenure_dupes %>%
        group_by(game_id, posteam, coach) %>%
        summarise(n = n(), .groups = 'drop') %>%
        count(n) %>%
        arrange(desc(n))
    )
}


  
  message("Step 2.7/7: Processing formation and defensive tendencies...")
  # Formation tendencies by down
  formation_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(posteam, down) %>%
    summarize(
      total_formation_plays = sum(!is.na(offense_formation)),
      
      # Formation usage rates
      shotgun_rate = ifelse(total_formation_plays > 0,
                           sum(offense_formation == "SHOTGUN", na.rm = TRUE) / total_formation_plays, 0),
      singleback_rate = ifelse(total_formation_plays > 0,
                              sum(offense_formation == "SINGLEBACK", na.rm = TRUE) / total_formation_plays, 0),
      empty_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "EMPTY", na.rm = TRUE) / total_formation_plays, 0),
      iform_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "I_FORM", na.rm = TRUE) / total_formation_plays, 0),
      pistol_rate = ifelse(total_formation_plays > 0,
                          sum(offense_formation == "PISTOL", na.rm = TRUE) / total_formation_plays, 0),
      wildcat_rate = ifelse(total_formation_plays > 0,
                           sum(offense_formation == "WILDCAT", na.rm = TRUE) / total_formation_plays, 0),
      jumbo_rate = ifelse(total_formation_plays > 0,
                         sum(offense_formation == "JUMBO", na.rm = TRUE) / total_formation_plays, 0),
      
      # Calculate success rates for each formation
      shotgun_success = mean(first_down[offense_formation == "SHOTGUN"], na.rm = TRUE),
      singleback_success = mean(first_down[offense_formation == "SINGLEBACK"], na.rm = TRUE),
      empty_success = mean(first_down[offense_formation == "EMPTY"], na.rm = TRUE),
      iform_success = mean(first_down[offense_formation == "I_FORM"], na.rm = TRUE),
      pistol_success = mean(first_down[offense_formation == "PISTOL"], na.rm = TRUE),
      wildcat_success = mean(first_down[offense_formation == "WILDCAT"], na.rm = TRUE),
      jumbo_success = mean(first_down[offense_formation == "JUMBO"], na.rm = TRUE),
      .groups = 'drop'
    )

  # Defense tendencies by down
  defense_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(defteam, down) %>%
    summarize(
      avg_pass_rushers = mean(number_of_pass_rushers, na.rm = TRUE),
      avg_box_defenders = mean(defenders_in_box, na.rm = TRUE),
      
      # Coverage type rates - using all known types
      total_coverage_plays = sum(!is.na(defense_coverage_type)),
      cover0_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_0", na.rm = TRUE) / total_coverage_plays, 0),
      cover1_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_1", na.rm = TRUE) / total_coverage_plays, 0),
      cover2_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_2", na.rm = TRUE) / total_coverage_plays, 0),
      cover3_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_3", na.rm = TRUE) / total_coverage_plays, 0),
      cover4_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_4", na.rm = TRUE) / total_coverage_plays, 0),
      cover6_rate = ifelse(total_coverage_plays > 0, 
                          sum(defense_coverage_type == "COVER_6", na.rm = TRUE) / total_coverage_plays, 0),
      two_man_rate = ifelse(total_coverage_plays > 0, 
                           sum(defense_coverage_type == "2_MAN", na.rm = TRUE) / total_coverage_plays, 0),
      prevent_rate = ifelse(total_coverage_plays > 0, 
                           sum(defense_coverage_type == "PREVENT", na.rm = TRUE) / total_coverage_plays, 0),
      .groups = 'drop'
    )

  # Coverage tendencies by down
  coverage_tendencies <- play_year %>%
    left_join(participation_data, by = c("game_id" = "nflverse_game_id", "play_id")) %>%
    filter(play_type %in% c("pass", "run")) %>%
    group_by(defteam, down, defense_coverage_type) %>%
    summarize(
      coverage_plays = n(),
      coverage_success_rate = mean(first_down == 0, na.rm = TRUE),  # defense success = stop
      .groups = 'drop'
    ) %>%
    group_by(defteam, down) %>%
    mutate(
      coverage_usage_rate = coverage_plays / sum(coverage_plays)
    ) %>%
    ungroup()
  
  
message("Step 3/7: Processing third and fourth downs...")
third_downs <- play_year %>%
    filter(down == 3) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type,
      vegas_wp,
      vegas_wpa,
      total_line,
      spread_line
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      season = year,
      # Adjust spread_line based on posteam
      spread_line = case_when(
        posteam == home_team ~ spread_line,
        posteam == away_team ~ -1 * spread_line,
        TRUE ~ NA_real_
      )
    ) %>%
    filter(!field_goal_attempt, !punt_attempt, play_type %in% c("pass", "run")) %>%
    mutate(converted = first_down) %>%
    select(-first_down) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id"))

fourth_downs <- play_year %>%
    filter(down == 4) %>%
    select(
      play_id,
      game_id,
      game_date,
      posteam,
      desc,
      home_team,
      away_team,
      ydstogo,
      yardline_100,
      game_seconds_remaining,
      posteam_timeouts_remaining,
      defteam_timeouts_remaining,
      posteam_score,
      defteam_score,
      play_type,
      yards_gained,
      first_down,
      rush_attempt,
      pass_attempt,
      field_goal_attempt,
      punt_attempt,
      home_coach,
      away_coach,
      week,
      roof,
      temp, 
      wind,
      drive,
      drive_start_transition,
      posteam_type,
      vegas_wp,
      vegas_wpa,
      total_line,
      spread_line
    ) %>%
    mutate(
      score_diff = posteam_score - defteam_score,
      posteam_coach = ifelse(posteam == home_team, home_coach, away_coach),
      attempt = if_else(play_type %in% c("punt", "field_goal"), 0, 1),
      season = year,
      # Adjust spread_line based on posteam
      spread_line = case_when(
        posteam == home_team ~ spread_line,
        posteam == away_team ~ -1 * spread_line,
        TRUE ~ NA_real_
      )
    ) %>%
    left_join(participation_data, 
              by = c("game_id" = "nflverse_game_id", "play_id" = "play_id"))

  message("Step 3.5/7: Assigning players to positions...")
  
  # Create empty player columns for third downs
  message("  Creating player columns for third downs...")
  for(i in 1:11) {
    third_downs[[paste0("offense_player_", i)]] <- NA_character_
    third_downs[[paste0("defense_player_", i)]] <- NA_character_
  }
  
  # Create empty player columns for fourth downs
  message("  Creating player columns for fourth downs...")
  for(i in 1:11) {
    fourth_downs[[paste0("offense_player_", i)]] <- NA_character_
    fourth_downs[[paste0("defense_player_", i)]] <- NA_character_
  }
  
# Updated position groupings with OTHER category
OFFENSE_POSITIONS <- list(
  ST = c("K", "P", "LS"),
  QB = "QB",
  BACKS = c("RB", "FB"),
  WR = "WR",
  TE = "TE", 
  OL = c("T", "G", "C", "OT", "OG"),
  OTHER = "OTHER"  # Catchall category
)

DEFENSE_POSITIONS <- list(
  ST = c("K", "P", "LS"),
  DL = c("DE", "DT", "NT", "DL"),
  LB = c("MLB", "ILB", "OLB", "LB"),
  DB = c("S", "FS", "SS", "SAF", "CB", "DB"),
  OTHER = "OTHER"  # Catchall category
)

assign_players_to_positions <- function(players_data, roster_data, is_offense = TRUE) {
  # Initialize empty slots
  player_slots <- rep(NA_character_, 11)
  next_slot <- 1
  
  # Skip if empty player data
  if(is.null(players_data) || players_data == "" || next_slot > 11) {
    return(player_slots)
  }
  
  # Get player ids from semicolon-separated string and count them
  player_ids <- unlist(strsplit(players_data, ";"))
  expected_players <- length(player_ids)
  
  if(expected_players == 0) {
    return(player_slots)
  }
  
  # Get player positions from roster data
  players_info <- roster_data %>%
    filter(gsis_id %in% player_ids) %>%
    select(gsis_id, depth_chart_position)
  
  if(nrow(players_info) == 0) {
    return(player_slots)
  }
  
  # Helper function to assign players of a position group
  assign_position_group <- function(position_group) {
    players <- players_info %>%
      filter(depth_chart_position %in% position_group)
    
    if(nrow(players) > 0 && next_slot <= 11) {
      slots_to_fill <- min(nrow(players), 12 - next_slot)
      player_slots[next_slot:(next_slot + slots_to_fill - 1)] <<- players$gsis_id[1:slots_to_fill]
      next_slot <<- next_slot + slots_to_fill
    }
  }
  
  # First handle special teams players
  assign_position_group(c("K", "P", "LS"))
  
  # Process regular position groups
  position_groups <- if(is_offense) OFFENSE_POSITIONS else DEFENSE_POSITIONS
  for(group in position_groups[!names(position_groups) %in% c("OTHER", "ST")]) {
    if(next_slot <= 11) {
      assign_position_group(group)
    } else {
      break
    }
  }
  
  # Fill any remaining slots with unassigned player IDs
  if(next_slot <= 11) {
    # Get all player IDs that haven't been assigned yet
    assigned_ids <- player_slots[!is.na(player_slots)]
    unassigned_ids <- setdiff(player_ids, assigned_ids)
    
    if(length(unassigned_ids) > 0) {
      slots_left <- 12 - next_slot
      ids_to_add <- min(length(unassigned_ids), slots_left)
      player_slots[next_slot:(next_slot + ids_to_add - 1)] <- unassigned_ids[1:ids_to_add]
    }
  }
  
  # Verify we've used as many players as possible
  actual_players <- sum(!is.na(player_slots))
  if(actual_players < min(expected_players, 11)) {
    # If we have fewer players than expected and slots available,
    # fill remaining slots with any unused IDs
    unused_ids <- setdiff(player_ids, player_slots[!is.na(player_slots)])
    remaining_slots <- which(is.na(player_slots))
    
    if(length(unused_ids) > 0 && length(remaining_slots) > 0) {
      slots_to_fill <- min(length(unused_ids), length(remaining_slots))
      player_slots[remaining_slots[1:slots_to_fill]] <- unused_ids[1:slots_to_fill]
    }
  }
  
  return(player_slots)
}
  
  # Apply position assignments to third downs
  message("  Processing third downs...")
  for(i in 1:nrow(third_downs)) {
    if(i %% 1000 == 0) message("    Processing row ", i, " of ", nrow(third_downs))
    
    # Process offense
    if(!is.na(third_downs$offense_players[i])) {
      player_slots <- assign_players_to_positions(third_downs$offense_players[i], roster_data, TRUE)
      for(j in 1:11) {
        third_downs[[paste0("offense_player_", j)]][i] <- player_slots[j]
      }
    }
    
    # Process defense
    if(!is.na(third_downs$defense_players[i])) {
      player_slots <- assign_players_to_positions(third_downs$defense_players[i], roster_data, FALSE)
      for(j in 1:11) {
        third_downs[[paste0("defense_player_", j)]][i] <- player_slots[j]
      }
    }
  }
  
  # Apply position assignments to fourth downs
  message("  Processing fourth downs...")
  for(i in 1:nrow(fourth_downs)) {
    if(i %% 1000 == 0) message("    Processing row ", i, " of ", nrow(fourth_downs))
    
    # Process offense
    if(!is.na(fourth_downs$offense_players[i])) {
      player_slots <- assign_players_to_positions(fourth_downs$offense_players[i], roster_data, TRUE)
      for(j in 1:11) {
        fourth_downs[[paste0("offense_player_", j)]][i] <- player_slots[j]
      }
    }
    
    # Process defense
    if(!is.na(fourth_downs$defense_players[i])) {
      player_slots <- assign_players_to_positions(fourth_downs$defense_players[i], roster_data, FALSE)
      for(j in 1:11) {
        fourth_downs[[paste0("defense_player_", j)]][i] <- player_slots[j]
      }
    }
  }
  
  message("Step 3.7/7: Counting original and assigned players...")
  
  # For third downs
  message("  Processing third downs...")
  third_downs <- third_downs %>%
    mutate(
      original_offense_count = ifelse(is.na(offense_players) | offense_players == "", 
                                    0, 
                                    lengths(strsplit(offense_players, ";"))),
      original_defense_count = ifelse(is.na(defense_players) | defense_players == "", 
                                    0, 
                                    lengths(strsplit(defense_players, ";")))
    ) %>%
    rowwise() %>%
    mutate(
      assigned_offense_count = sum(!is.na(c_across(starts_with("offense_player_")))),
      assigned_defense_count = sum(!is.na(c_across(starts_with("defense_player_"))))
    ) %>%
    ungroup()
  
  # For fourth downs
  message("  Processing fourth downs...")
  fourth_downs <- fourth_downs %>%
    mutate(
      original_offense_count = ifelse(is.na(offense_players) | offense_players == "", 
                                    0, 
                                    lengths(strsplit(offense_players, ";"))),
      original_defense_count = ifelse(is.na(defense_players) | defense_players == "", 
                                    0, 
                                    lengths(strsplit(defense_players, ";")))
    ) %>%
    rowwise() %>%
    mutate(
      assigned_offense_count = sum(!is.na(c_across(starts_with("offense_player_")))),
      assigned_defense_count = sum(!is.na(c_across(starts_with("defense_player_"))))
    ) %>%
    ungroup()

  message("Step 4/7: Adding player information...")
  for(i in 1:11) {
    cat("\rProcessing player", i, "of 11")
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("offense_player_", i), 
                                  paste0("offense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("offense_player_", i), 
                                   paste0("offense_player_", i))
    
    third_downs <- add_player_info(third_downs, 
                                  roster_data, 
                                  paste0("defense_player_", i), 
                                  paste0("defense_player_", i))
    fourth_downs <- add_player_info(fourth_downs, 
                                   roster_data, 
                                   paste0("defense_player_", i), 
                                   paste0("defense_player_", i))
  }
  
  message("Step 5/7: Processing team injury data...")
  third_downs <- third_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  fourth_downs <- fourth_downs %>%
    mutate(defteam = ifelse(posteam == home_team, away_team, home_team)) %>%
    left_join(run_pcts %>% 
                mutate(join_week = week + 1) %>% 
                select(-week), 
              by = c("posteam", "week" = "join_week")) %>%
    left_join(prep_time_df %>% 
                select(posteam, game_date, prep_days),
              by = c("posteam", "game_date"))

  third_downs <- third_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  fourth_downs <- fourth_downs %>%
    rowwise() %>%
    mutate(
      offense_injuries = list(get_team_injury_counts(posteam, game_date)),
      defense_injuries = list(get_team_injury_counts(defteam, game_date))
    ) %>%
    ungroup() %>%
    unnest_wider(c(offense_injuries, defense_injuries), 
                 names_sep = "_")

  third_downs <- third_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  fourth_downs <- fourth_downs %>%
    group_by(posteam, game_date) %>%
    mutate(
      offense_avg_days_since_injury = mean(c_across(matches("offense_player_\\d+_days_since_injury")), na.rm = TRUE),
      offense_total_recent_injuries = sum(c_across(matches("offense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    group_by(defteam, game_date) %>%
    mutate(
      defense_avg_days_since_injury = mean(c_across(matches("defense_player_\\d+_days_since_injury")), na.rm = TRUE),
      defense_total_recent_injuries = sum(c_across(matches("defense_player_\\d+_injury_count_1yr")), na.rm = TRUE)
    ) %>%
    ungroup()

  all_teams_weeks <- bind_rows(
    select(third_downs, defteam, week),
    select(fourth_downs, defteam, week)
  ) %>% distinct()

  prev_stats <- map2_dfr(
    all_teams_weeks$defteam,
    all_teams_weeks$week,
    ~{
      prev_game <- play_year %>%
        filter((home_team == .x | away_team == .x) & week < .y) %>%
        arrange(desc(week)) %>%
        slice(1) %>%
        pull(week)
      
      if(length(prev_game) == 0) return(tibble(
        defteam = .x,
        week = .y,
        attempts = 0,
        successes = 0,
        failures = 0
      ))
      
      play_year %>%
        filter(week == prev_game,
               (home_team == .x | away_team == .x),
               down == 4,
               !play_type %in% c("punt", "field_goal")) %>%
        summarize(
          defteam = .x,
          week = .y,
          attempts = n(),
          successes = sum(first_down == 1, na.rm = TRUE),
          failures = sum(first_down == 0, na.rm = TRUE)
        )
    })

  message("Step 5.9/7: Debugging data structures section...")


# Add fixed down values since all third downs have down=3
third_downs$down <- 3
fourth_downs$down <- 4


  
message("Step 6 Diagnostics: Checking key data structures and columns...")



  model_3rd <- third_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  model_4th <- fourth_downs %>%
    mutate(
      play_type_clean = case_when(
        pass_attempt == 1 ~ "pass",
        rush_attempt == 1 ~ "rush",
        TRUE ~ "other"
      )
    ) %>%
    left_join(prev_stats, by = c("defteam", "week")) %>%
    group_by(posteam) %>%
    arrange(week) %>%
    mutate(
      down1_pct = zoo::na.locf(down1_pct, na.rm = FALSE),
      down2_pct = zoo::na.locf(down2_pct, na.rm = FALSE),
      down3_pct = zoo::na.locf(down3_pct, na.rm = FALSE)
    ) %>%
    ungroup()

  
  # Add before returning the models in process_year:
  model_3rd <- model_3rd %>%
    left_join(coach_records, by = c("posteam_coach" = "coach")) %>%
    left_join(current_team_records, by = c("posteam_coach" = "coach", "posteam")) %>%
    left_join(game_tenure %>% 
                select(game_id, posteam, coach, tenure_days), 
              by = c("game_id", "posteam", "posteam_coach" = "coach")) %>%
    left_join(formation_tendencies %>% 
                filter(down == 3), 
              by = c("posteam", "down")) %>%
    left_join(defense_tendencies %>% 
                filter(down == 3), 
              by = c("defteam", "down")) %>%
    left_join(coverage_tendencies %>% 
                filter(down == 3), 
              by = c("defteam", "defense_coverage_type"))

  model_4th <- model_4th %>%
    left_join(coach_records, by = c("posteam_coach" = "coach")) %>%
    left_join(current_team_records, by = c("posteam_coach" = "coach", "posteam")) %>%
    left_join(game_tenure %>% 
                select(game_id, posteam, coach, tenure_days), 
              by = c("game_id", "posteam", "posteam_coach" = "coach")) %>%
    left_join(formation_tendencies %>% 
                filter(down == 4), 
              by = c("posteam", "down")) %>%
    left_join(defense_tendencies %>% 
                filter(down == 4), 
              by = c("defteam", "down")) %>%
    left_join(coverage_tendencies %>% 
                filter(down == 4), 
              by = c("defteam", "defense_coverage_type"))
  
  # Add after final model joins in Step 6:
message("Diagnosing final model duplicates...")
model_4th_dupes <- model_4th %>%
  group_by(play_id, game_id) %>%
  filter(n() > 1)

print("Number of duplicate rows in final fourth down model:")
print(nrow(model_4th_dupes))
if(nrow(model_4th_dupes) > 0) {
  print("Sample of duplicates with key columns:")
  print(model_4th_dupes %>% 
        select(play_id, game_id, posteam, defteam, offense_formation, defense_coverage_type) %>%
        head())
}
  
  
  message("Step 7/7: Adding specialists...")
  specialists <- get_specialists(year)

  specialists$kickers <- specialists$kickers %>%
    rename(
      kicker_id = kicker_player_id,
      kicker_name = kicker_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, kicker_id, kicker_name)

  specialists$punters <- specialists$punters %>%
    rename(
      punter_id = punter_player_id,
      punter_name = punter_player_name
    ) %>%
    filter(season == year | season == year - 1) %>%
    arrange(desc(season), desc(week)) %>%
    distinct(team, .keep_all = TRUE) %>%
    select(team, punter_id, punter_name)

  model_4th <- model_4th %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  model_3rd <- model_3rd %>%
    left_join(specialists$kickers, by = c("posteam" = "team")) %>%
    left_join(specialists$punters, by = c("posteam" = "team"))

  return(list(
    third_downs = model_3rd,
    fourth_downs = model_4th
  ))
}
```

```{r}
# Main processing
years <- 2017#2016:2023
all_data <- list()

for(year in years) {
  all_data[[as.character(year)]] <- process_year(year)
}

combined_third_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$third_downs))
)

combined_fourth_downs <- bind_rows(
  lapply(all_data, function(x) validate_data(x$fourth_downs))
)
```



# filter out weeks with no info on who was on the field

get the rows in cfd and ctd that have a na in either offense_players or defense_players
```{r}
# Function to get max regular season week based on year
get_max_regular_season_week <- function(year) {
  if (year >= 2021) {
    return(18)  # 17-game season starting in 2021
  } else {
    return(17)  # 16-game season before 2021
  }
}

# Filter out playoff games from combined datasets
combined_third_downs <- combined_third_downs %>%
  filter(
    (season < 2021 & week <= 17) |  # Pre-2021 seasons
    (season >= 2021 & week <= 18)    # 2021 and later seasons
  )

combined_fourth_downs <- combined_fourth_downs %>%
  filter(
    (season < 2021 & week <= 17) |  # Pre-2021 seasons
    (season >= 2021 & week <= 18)    # 2021 and later seasons
  )

# Verify the filtering worked
print("Third downs week range by season:")
print(combined_third_downs %>% 
  group_by(season) %>% 
  summarise(min_week = min(week), max_week = max(week)))

print("\nFourth downs week range by season:")
print(combined_fourth_downs %>% 
  group_by(season) %>% 
  summarise(min_week = min(week), max_week = max(week)))
```



```{r}
# Get counts before filtering by season and week
before_counts_third <- combined_third_downs %>%
  group_by(season, week) %>%
  summarise(count_before = n())

before_counts_fourth <- combined_fourth_downs %>%
  group_by(season, week) %>%
  summarise(count_before = n())

# Filter and get counts after by season and week
after_counts_third <- combined_third_downs %>%
  filter(!is.na(offense_players) & !is.na(defense_players)) %>%
  group_by(season, week) %>%
  summarise(count_after = n())

after_counts_fourth <- combined_fourth_downs %>%
  filter(!is.na(offense_players) & !is.na(defense_players)) %>%
  group_by(season, week) %>%
  summarise(count_after = n())

# Compare before and after for third downs
lost_rows_third <- before_counts_third %>%
  left_join(after_counts_third, by = c("season", "week")) %>%
  mutate(rows_lost = count_before - count_after) %>%
  filter(rows_lost > 0) %>%
  arrange(season, week)

# Compare before and after for fourth downs
lost_rows_fourth <- before_counts_fourth %>%
  left_join(after_counts_fourth, by = c("season", "week")) %>%
  mutate(rows_lost = count_before - count_after) %>%
  filter(rows_lost > 0) %>%
  arrange(season, week)

# Print results
cat("Third Downs - Lost Rows by Season and Week:\n")
print(lost_rows_third)

cat("\nFourth Downs - Lost Rows by Season and Week:\n")
print(lost_rows_fourth)

# Print totals
cat("\nTotal rows removed from Third Downs:", sum(lost_rows_third$rows_lost))
cat("\nTotal rows removed from Fourth Downs:", sum(lost_rows_fourth$rows_lost))



# Filter out rows with missing player information
combined_third_downs <- combined_third_downs %>%
  filter(!is.na(offense_players) & !is.na(defense_players))

combined_fourth_downs <- combined_fourth_downs %>%
  filter(!is.na(offense_players) & !is.na(defense_players))
```




```{r}
# Read the attendance data
attendance_data <- read.csv("nfl_attendance_2016_2023.csv")
# Revised add_attendance function
add_attendance <- function(pbp_df, attendance_df) {
  pbp_df %>%
    left_join(
      # Reshape attendance data to long format for joining
      attendance_df %>%
        pivot_longer(
          cols = starts_with("Week."),  # Match Week. columns
          names_to = "week_label",
          values_to = "attendance"
        ) %>%
        # Extract week number, accounting for the dot
        mutate(
          week = as.numeric(gsub("Week\\.", "", week_label)),
          home_attendance = attendance
        ) %>%
        select(Team, Year, week, home_attendance),
      # Join conditions
      by = c(
        "home_team" = "Team",
        "season" = "Year",
        "week" = "week"
      )
    )
}

# Apply to your pbp datasets
combined_third_downs <- add_attendance(combined_third_downs, attendance_data)
combined_fourth_downs <- add_attendance(combined_fourth_downs, attendance_data)

# Verify the join worked
print("Number of NA attendance values in third downs:")
print(sum(is.na(combined_third_downs$home_attendance)))
print("\nNumber of NA attendance values in fourth downs:")
print(sum(is.na(combined_fourth_downs$home_attendance)))
```


```{r}
library(nflreadr)
library(tidyverse)
library(slider)

# Create season-aware team mapping function
team_name_mapping <- function(team, season) {
  case_when(
    # Chargers: SD -> LAC in 2017
    team == "LAC" & season < 2017 ~ "SD",
    team == "SD" & season >= 2017 ~ "LAC",
    # Rams: STL/SL -> LA in 2016
    team %in% c("LA", "SL") & season < 2016 ~ "STL",
    team == "STL" & season >= 2016 ~ "LA",
    # Raiders: OAK -> LV in 2020
    team == "LV" & season < 2020 ~ "OAK",
    team == "OAK" & season >= 2020 ~ "LV",
    # Other historical fixes
    team == "ARZ" ~ "ARI",
    team == "BLT" ~ "BAL",
    team == "CLV" ~ "CLE",
    team == "HST" ~ "HOU",
    # If no mapping needed, return original
    TRUE ~ team
  )
}

# Process roster data for team changes
player_team_changes <- load_rosters_weekly(seasons = 2015:2023) %>%
  filter(
    (season < 2021 & week <= 17) |  
    (season >= 2021 & week <= 18)    
  ) %>%
  mutate(
    team = map2_chr(team, season, team_name_mapping)
  ) %>%
  select(player_id = gsis_id, team, season, week) %>%
  group_by(player_id, season, week) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  arrange(player_id, season, week) %>%
  group_by(player_id) %>%
  mutate(
    prev_team = lag(team),
    prev_season = lag(season),
    prev_week = lag(week)
  ) %>%
  filter(team != prev_team, !is.na(prev_team)) %>%
  # Remove relocations
  filter(!(
    # Rams relocation
    (prev_team == "STL" & team == "LA" & prev_season == 2015 & season == 2016) |
    # Chargers relocation
    (prev_team == "SD" & team == "LAC" & prev_season == 2016 & season == 2017) |
    # Raiders relocation
    (prev_team == "OAK" & team == "LV" & prev_season == 2019 & season == 2020)
  )) %>%
  select(
    player_id,
    season_from = prev_season,
    week_from = prev_week,
    season_to = season,
    week_to = week,
    old_team = prev_team,
    new_team = team
  ) %>%
  ungroup()

# Load and process games data
games_data <- load_schedules(seasons = 2013:2023) %>%
  filter(game_type == "REG") %>%
  mutate(
    home_team = map2_chr(home_team, season, team_name_mapping),
    away_team = map2_chr(away_team, season, team_name_mapping)
  )

# Create team-level data with wins
team_games <- bind_rows(
  # Home games
  games_data %>%
    select(season, week, team = home_team, opponent = away_team, 
           team_score = home_score, opp_score = away_score) %>%
    mutate(win = case_when(
      team_score > opp_score ~ 1,
      team_score < opp_score ~ 0,
      TRUE ~ 0.5  # ties count as half win
    )),
  # Away games  
  games_data %>%
    select(season, week, team = away_team, opponent = home_team, 
           team_score = away_score, opp_score = home_score) %>%
    mutate(win = case_when(
      team_score > opp_score ~ 1,
      team_score < opp_score ~ 0,
      TRUE ~ 0.5
    ))
) %>%
  arrange(season, week, team)

# Calculate rolling wins for each team
team_rolling_wins <- team_games %>%
  group_by(team) %>%
  arrange(season, week) %>%
  mutate(
    wins_last_18 = slide_sum(win, before = 17, after = 0)
  ) %>%
  ungroup()

# Final join
player_team_changes_with_wins <- player_team_changes %>%
  left_join(
    team_rolling_wins %>% 
      select(season, week, team, wins_last_18),
    by = c(
      "season_from" = "season",
      "week_from" = "week",
      "old_team" = "team"
    )
  ) %>%
  # Optionally rename the wins column to be more descriptive
  rename(old_team_wins_last_18 = wins_last_18)
```

```{r}
library(zoo)
# Create a complete season-week grid for each player
player_weeks <- load_rosters_weekly(seasons = 2015:2023) %>%
  select(player_id = gsis_id, season) %>%
  distinct() %>%
  # Create all possible weeks for each player-season
  crossing(
    week = 1:max(
      if_else(season < 2021, 17, 18)
    )
  ) %>%
  arrange(player_id, season, week)

# Join with roster data to get team information
player_status <- load_rosters_weekly(seasons = 2015:2023) %>%
  filter(
    (season < 2021 & week <= 17) |  
    (season >= 2021 & week <= 18)    
  ) %>%
  mutate(
    team = map2_chr(team, season, team_name_mapping)
  ) %>%
  select(player_id = gsis_id, team, season, week) %>%
  # Get latest team status for each week
  group_by(player_id, season, week) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  right_join(player_weeks, by = c("player_id", "season", "week")) %>%
  arrange(player_id, season, week) %>%
  group_by(player_id) %>%
  # Fill in team status for weeks between changes
  fill(team, .direction = "down") %>%
  # Create sequential week number starting from 2015
  mutate(
    sequential_week = case_when(
      season == 2015 ~ week,
      season == 2016 ~ week + 17,
      season == 2017 ~ week + 34,
      season == 2018 ~ week + 51,
      season == 2019 ~ week + 68,
      season == 2020 ~ week + 85,
      season == 2021 ~ week + 102,
      season == 2022 ~ week + 120,
      season == 2023 ~ week + 138
    ),
    prev_team = lag(team),
    team_switch = team != prev_team & !is.na(prev_team),
    # Find the most recent switch week for each row
    last_switch_week = lag(if_else(team_switch, sequential_week, NA_real_)),
    last_switch_week = na.locf(last_switch_week, na.rm = FALSE),
    # Calculate weeks since most recent switch
    weeks_since_switch = if_else(
      is.na(last_switch_week),
      0,  # For players with no prior switches
      sequential_week - last_switch_week
    )
  ) %>%
  ungroup()

# Join with team performance data
final_player_status <- player_status %>%
  left_join(
    team_rolling_wins %>% 
      select(season, week, team, wins_last_18),
    by = c("season", "week", "team")
  ) %>%
  group_by(player_id) %>%
  mutate(
    # Create temp columns for team switch info
    temp_old_team = if_else(team_switch, prev_team, NA_character_),
    temp_old_wins = if_else(team_switch, lag(wins_last_18), NA_real_),
    # Carry forward these values until next switch
    old_team = na.locf(temp_old_team, na.rm = FALSE),
    old_team_wins = na.locf(temp_old_wins, na.rm = FALSE)
  ) %>%
  ungroup() %>%
  # Clean up and select final columns
  select(
    player_id,
    season,
    week,
    current_team = team,
    old_team,
    team_switch,
    weeks_since_switch,
    current_team_wins = wins_last_18,
    old_team_wins
  )
```


```{r}
library(rlang)

add_switch_info <- function(pbp_data, switch_data) {
  # For each player column (offense 1-11, defense 1-11, kicker, punter)
  result <- pbp_data
  
  # Add kicker and punter info
  if("kicker_id" %in% names(result)) {
    result <- result %>%
      left_join(
        switch_data %>% 
          select(player_id, season, week, weeks_since_switch, old_team, old_team_wins),
        by = c("kicker_id" = "player_id", "season", "week"),
        suffix = c("", "_kicker")
      )
  }
  
  if("punter_id" %in% names(result)) {
    result <- result %>%
      left_join(
        switch_data %>% 
          select(player_id, season, week, weeks_since_switch, old_team, old_team_wins),
        by = c("punter_id" = "player_id", "season", "week"),
        suffix = c("", "_punter")
      )
  }
  
  # Process offensive players
  for(i in 1:11) {
    player_col <- paste0("offense_player_", i)
    join_by <- setNames(c("player_id", "season", "week"), c(player_col, "season", "week"))
    result <- result %>%
      left_join(
        switch_data %>% 
          select(player_id, season, week, weeks_since_switch, old_team, old_team_wins),
        by = join_by,
        suffix = c("", paste0("_o", i))
      )
  }
  
  # Process defensive players
  for(i in 1:11) {
    player_col <- paste0("defense_player_", i)
    join_by <- setNames(c("player_id", "season", "week"), c(player_col, "season", "week"))
    result <- result %>%
      left_join(
        switch_data %>% 
          select(player_id, season, week, weeks_since_switch, old_team, old_team_wins),
        by = join_by,
        suffix = c("", paste0("_d", i))
      )
  }
  
  return(result)
}

# Apply to both datasets
combined_third_downs<- add_switch_info(combined_third_downs, final_player_status)
combined_fourth_downs <- add_switch_info(combined_fourth_downs, final_player_status)
```

```{r}
# Save results
write.csv(combined_third_downs, "ctd.csv", row.names = FALSE)
write.csv(combined_fourth_downs, "cfd.csv", row.names = FALSE)
```




```{r}
library(readr)
library(dplyr)
combined_third_downs <- read.csv("ctd.csv")
combined_fourth_downs <- read.csv("cfd.csv")
```



# Filtering for only 4th down attempts that are a rush_attempt = 1 or pass_attempt = 1
```{r}
attempts_4th <- combined_fourth_downs %>%
  filter(rush_attempt == 1 | pass_attempt == 1)
```

```{r}
# Define defensive positions
defensive_positions <- c("K", "P", "LS", "DB", "CB", "S", "FS", "SS", "LB", "ILB", "OLB", "MLB", "DL", "DE", "DT", "NT", "EDGE")

# Function to check if any offensive player has a defensive position
check_defensive_positions <- function(row) {
  # Get all offensive player positions
  positions <- c()
  for(i in 1:11) {
    pos_col <- paste0("offense_player_", i, "_position")
    if(pos_col %in% names(row)) {
      positions <- c(positions, row[[pos_col]])
    }
  }
  
  # Return TRUE if no defensive positions found
  !any(positions %in% defensive_positions, na.rm = TRUE)
}

# Apply the filter
clean_attempts_4th <- attempts_4th %>%
  filter(mapply(check_defensive_positions, split(., 1:nrow(.))))

# Print the number of rows removed
cat("Original rows:", nrow(attempts_4th), "\n")
cat("Rows after filtering:", nrow(clean_attempts_4th), "\n")
cat("Rows removed:", nrow(attempts_4th) - nrow(clean_attempts_4th), "\n")

# Optional: Examine some examples of removed plays
removed_plays <- attempts_4th %>%
  filter(!mapply(check_defensive_positions, split(., 1:nrow(.)))) %>%
  select(play_id, game_id, desc, matches("offense_player_\\d+_position"))

# Show first few removed plays
head(removed_plays)
```


check in the desc column if the word put or field goal insensitive to caps is present then kill it

```{r}
# Filter out plays with "punt" or "field goal" in the description
clean_attempts_4th <- clean_attempts_4th %>%
  filter(!grepl("punt", desc, ignore.case = TRUE) & !grepl("field goal", desc, ignore.case = TRUE))
```

write to cafd
```{r}
write.csv(clean_attempts_4th, "cafd.csv", row.names = FALSE)
```


# the gsis id problem

get the players and k and p gsis ids then the offense and defense id cols so 26 cols in total into there own df with desc so 27 cols.. we will use 3rd down for just now

```{r}
# Read the combined third downs data
combined_third_downs <- read.csv("ctd.csv")

# Create vector of column names we want to keep
id_columns <- c(
  "game_id",
  "desc",
  "offense_players",
  "defense_players",
  "kicker_id",
  "punter_id",
  paste0("offense_player_", 1:11),
  paste0("defense_player_", 1:11)
)

# Create the check_ids dataframe
check_ids <- combined_third_downs[, id_columns]

# Print some summary information
cat("Dimensions of check_ids:", dim(check_ids), "\n")
cat("Column names:", colnames(check_ids), "\n")

# Check for NA values in each column
na_counts <- sapply(check_ids, function(x) sum(is.na(x)))
print("NA counts per column:")
print(na_counts)
```


now count how many values offense_players and defense_players have in them

```{r}
str(check_ids)

```


```{r}
check_ids <- check_ids %>%
  mutate(
    off_count = sapply(strsplit(offense_players, ";"), length),
    def_count = sapply(strsplit(defense_players, ";"), length)
  )
```

the 12s on def are heavily carolina

THE TUSH PUSH EAGLES NEED TO DIE




