---
title: "PFF Variable Analysis 2015"
output: html_document
date: "2024-12-31"
---

```{r}
library(readr)
library(dplyr)
library(tidyr)
```

# Passing Reports
## Passing Grades
```{r}
passing_grades <- read_csv("yearly_data/2015/passing_summary.csv")
print("Passing Grades Columns:")
print(colnames(passing_grades))
```
## Passing Depth
```{r}
passing_depth <- read_csv("yearly_data/2015/passing_depth.csv")
print("Passing Depth Columns:")
print(colnames(passing_depth))
```
## Passing Pressure
```{r}
passing_pressure <- read_csv("yearly_data/2015/passing_pressure.csv")
print("Passing Pressure Columns:")
print(colnames(passing_pressure))
```
## Passing Concept
```{r}
passing_concept <- read_csv("yearly_data/2015/passing_concept.csv")
print("Passing Concept Columns:")
print(colnames(passing_concept))
```
## Time In Pocket
```{r}
time_in_pocket <- read_csv("yearly_data/2015/time_in_pocket.csv")
print("Time In Pocket Columns:")
print(colnames(time_in_pocket))
```
## Allowed Pressure
```{r}
allowed_pressure <- read_csv("yearly_data/2015/passing_allowed_pressure.csv")
print("Allowed Pressure Columns:")
print(colnames(allowed_pressure))
```
# Receiving Reports
## Receiving Grades
```{r}
receiving_grades <- read_csv("yearly_data/2015/receiving_summary.csv")
print("Receiving Grades Columns:")
print(colnames(receiving_grades))
```
## Receiving Depth
```{r}
receiving_depth <- read_csv("yearly_data/2015/receiving_depth.csv")
print("Receiving Depth Columns:")
print(colnames(receiving_depth))
```
## Receiving Concept
```{r}
receiving_concept <- read_csv("yearly_data/2015/receiving_concept.csv")
print("Receiving Concept Columns:")
print(colnames(receiving_concept))
```
## Receiving vs Scheme
```{r}
receiving_vs_scheme <- read_csv("yearly_data/2015/receiving_scheme.csv")
print("Receiving vs Scheme Columns:")
print(colnames(receiving_vs_scheme))
```
# Rushing Reports
## Rushing Grades
```{r}
rushing_grades <- read_csv("yearly_data/2015/rushing_summary.csv")
print("Rushing Grades Columns:")
print(colnames(rushing_grades))
```
# Blocking Reports
## Blocking Grades
```{r}
blocking_grades <- read_csv("yearly_data/2015/offense_blocking.csv")
print("Blocking Grades Columns:")
print(colnames(blocking_grades))
```
## Pass Blocking
```{r}
pass_blocking <- read_csv("yearly_data/2015/offense_pass_blocking.csv")
print("Pass Blocking Columns:")
print(colnames(pass_blocking))
```
## Run Blocking
```{r}
run_blocking <- read_csv("yearly_data/2015/offense_run_blockng.csv")
print("Run Blocking Columns:")
print(colnames(run_blocking))
```
# Defense Reports
## Defense Grades
```{r}
defensive_grades <- read_csv("yearly_data/2015/defense_summary.csv")
print("Defense Grades Columns:")
print(colnames(defensive_grades))
```
## Pass Rush Grades
```{r}
pass_rush_grades <- read_csv("yearly_data/2015/pass_rush_summary.csv")
print("Pass Rush Grades Columns:")
print(colnames(pass_rush_grades))
```
## Run Defense Grades
```{r}
run_defense_grades <- read_csv("yearly_data/2015/run_defense_summary.csv")
print("Run Defense Grades Columns:")
print(colnames(run_defense_grades))
```
## Coverage Grades
```{r}
coverage_grades <- read_csv("yearly_data/2015/defense_coverage_summary.csv")
print("Coverage Grades Columns:")
print(colnames(coverage_grades))
```
## Coverage Scheme
```{r}
coverage_schemes <- read_csv("yearly_data/2015/defense_coverage_scheme.csv")
print("Coverage Scheme Columns:")
print(colnames(coverage_schemes))
```
## Slot Coverage
```{r}
slot_coverage <- read_csv("yearly_data/2015/slot_coverage.csv")
print("Slot Coverage Columns:")
print(colnames(slot_coverage))
```
## Pass Rushing Productivity
```{r}
pass_rushing_productivity <- read_csv("yearly_data/2015/pass_rush_productivity.csv")
print("Pass Rushing Productivity Columns:")
print(colnames(pass_rushing_productivity))
```
# Special Teams Reports
## Special Teams Grades
```{r}
special_teams_grades <- read_csv("yearly_data/2015/special_teams_summary.csv")
print("Special Teams Grades Columns:")
print(colnames(special_teams_grades))
```
## Field Goal Grades
```{r}
field_goals_grades <- read_csv("yearly_data/2015/field_goal_summary.csv")
print("Field Goal Grades Columns:")
print(colnames(field_goals_grades))
```
## Kickoff Grades
```{r}
kickoff_grades <- read_csv("yearly_data/2015/kickoff_summary.csv")
print("Kickoff Grades Columns:")
print(colnames(kickoff_grades))
```
## Punting Grades
```{r}
punting_grades <- read_csv("yearly_data/2015/punting_summary.csv")
print("Punting Grades Columns:")
print(colnames(punting_grades))
```
## Kickoff and Punt Return Grades
```{r}
kickoff_and_punt_return_grades <- read_csv("yearly_data/2015/return_summary.csv")
print("Kickoff and Punt Return Grades Columns:")
print(colnames(kickoff_and_punt_return_grades))
```

# Column Analysis Summary
```{r}
# Create list of all dataframes with their names
all_dfs <- list(
  "passing_grades" = passing_grades,
  "passing_depth" = passing_depth,
  "passing_pressure" = passing_pressure,
  "passing_concept" = passing_concept,
  "time_in_pocket" = time_in_pocket,
  "allowed_pressure" = allowed_pressure,
  "receiving_grades" = receiving_grades,
  "receiving_depth" = receiving_depth,
  "receiving_concept" = receiving_concept,
  "receiving_vs_scheme" = receiving_vs_scheme,
  "rushing_grades" = rushing_grades,
  "blocking_grades" = blocking_grades,
  "pass_blocking" = pass_blocking,
  "run_blocking" = run_blocking,
  "defensive_grades" = defensive_grades,
  "pass_rush_grades" = pass_rush_grades,
  "run_defense_grades" = run_defense_grades,
  "coverage_grades" = coverage_grades,
  "coverage_schemes" = coverage_schemes,
  "slot_coverage" = slot_coverage,
  "pass_rushing_productivity" = pass_rushing_productivity,
  "special_teams_grades" = special_teams_grades,
  "field_goals_grades" = field_goals_grades,
  "kickoff_grades" = kickoff_grades,
  "punting_grades" = punting_grades,
  "kickoff_and_punt_return_grades" = kickoff_and_punt_return_grades
)

# Create a dataframe showing which columns appear in which reports
column_summary <- data.frame(
  column_name = character(),
  appears_in = character(),
  num_reports = numeric(),
  stringsAsFactors = FALSE
)

# For each unique column name across all dataframes
unique_cols <- unique(unlist(lapply(all_dfs, colnames)))

for(col in unique_cols) {
  # Find which reports contain this column
  reports_with_col <- names(all_dfs)[sapply(all_dfs, function(df) col %in% colnames(df))]
  
  # Add to summary dataframe
  column_summary <- rbind(column_summary, 
                         data.frame(
                           column_name = col,
                           appears_in = paste(reports_with_col, collapse = ", "),
                           num_reports = length(reports_with_col),
                           stringsAsFactors = FALSE
                         ))
}

# Sort by number of reports (descending) and alphabetically by column name
column_summary <- column_summary %>%
  arrange(desc(num_reports), column_name)

# Print the final summary
print("\nColumn Summary:")
print(column_summary, row.names = FALSE)
```


